---
title: "GB1, Ostrov codes, random walk"
author: "Hana Rozhonova"
date: "7/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(stringr)
library(ggplot2)
library(gridExtra)
library(ggpubr)
library("viridis")
library(data.table)
```

# Read the data
```{r}
if (file.exists("output/results_10000/results_processed.tsv")) {
  results = read.delim("output/results_10000/results_processed.tsv", header=TRUE)
  results$numSplitBlocks = as.factor(results$numSplitBlocks)
  results$numStop = as.factor(results$numStop)
} else {
  results = read.delim("output/results_10000/results", header=FALSE)
  colnames(results) = c("Id", "Block1", "Block2", "Block3", "Block4", "X", "Z", "B", "J", "Robustness", paste0("Fitness_", seq(0,990,by=10)))
   # number of split codon blocks
  results$numSplitBlocks = as.factor(apply(results, 1, function(x) sum(x[c("X", "Z", "B", "J")]!="-")))
  # number of stop codons
  results$numStop = as.factor(3 - (results$Block2!="-") + 2*apply(results, 1, function(x) sum(x[c("Block1", "Block3", "Block4")]=="O")))
  
  write.table(results, file="output/results_10000/results_processed.tsv", quote=FALSE, sep="\t", row.names=FALSE)
}
```

# Correlations between code robustness and mean fitness
Correlation test
```{r}
cor.test(results$Robustness, results$Fitness_500)
```

Restrict to only those codes where the size of the summit is the same as in the SGC and the summit is not split:
```{r}
results$nCodons_W = 1+2*(results$Block1=="W") + 1*(results$Block2=="W") + 2*(results$Block3=="W") + 2*(results$Block4=="W")
results$nCodons_L = 6-2*(results$Block1!="-")+1*(results$Block2=="L") 
results$nCodons_A = 4+2*(results$Block3=="A") + 2*(results$Block4=="A")
results$SummitSize = (results$nCodons_W**2) * results$nCodons_A * results$nCodons_L

results$splitW = rowSums(results[,6:9]=="W" )
results$splitL = rowSums(results[,6:9]=="L")
results$splitA = rowSums(results[,6:9]=="A")

tmp_results = results[results$SummitSize==results$SummitSize[results$Id=="standard"] &
                results$splitA==0 & results$splitL==0 & results$splitW==0,]

cor.test(tmp_results$Fitness_500, tmp_results$Robustness)
```

# Effect of number of split codon blocks
```{r}
means = c()
maxs = c()
mins = c()
steps = c()
numSplits = c()
for (numSplit in unique(results$numSplitBlocks)) {
  tmp = results[results$numSplitBlocks==numSplit,]
  for (i in seq(0,990,10)) {
    m = mean(tmp[,(i/10)+11])
    means = c(means, m)
    sd = sd(tmp[,(i/10)+11])
    mins = c(mins, m-sd)
    maxs = c(maxs, m+sd)
    steps = c(steps, i)
  }
  numSplits = c(numSplits, rep(numSplit, 100))
}
data = data.frame(numSteps = steps,
                  mean = means,
                  min = mins,
                  max = maxs,
                  numSplit = numSplits)
data$numSplit = as.factor(data$numSplit)

ggplot(data, aes(x=numSteps, y=mean, color=numSplit, fill=numSplit)) + 
  geom_line(size=0.4) + 
  #geom_ribbon(aes(ymin=data$min, ymax=data$max), alpha=0.1) +
  scale_color_viridis_d() + scale_fill_viridis_d() +
  theme_test()

ggplot(data, aes(x=numSteps, y=mean, color=numSplit, fill=numSplit)) + 
  geom_line(size=0.4) + 
  #geom_ribbon(aes(ymin=data$min, ymax=data$max), alpha=0.1) +
  scale_color_viridis_d() + scale_fill_viridis_d() +
  theme_test() +
  xlim(c(100,250)) + ylim(0.2,0.8)
```

Correlation between number of split codon blocks and fitness:
```{r}
cors = c()
pvals = c()
for (i in 11:110) {
  test = cor.test(as.numeric(results$numSplitBlocks), results[,i])
  cors = c(cors, test$estimate)
  pvals = c(pvals, test$p.value)
}
data = data.frame(numSteps = seq(0,990,10),
                  cor = cors,
                  pval = pvals)
(p_cor = ggplot(data, aes(x=numSteps, y=cor)) +
  geom_point() +
  geom_line() +
  theme_test())

p_pval = ggplot(data, aes(x=numSteps, y=pval)) +
  geom_point() +
  geom_line() +
  theme_test()

grid.arrange(p_cor, p_pval, ncol=2)
```

Only codes that preserve the size of the global peak:
```{r}
cor.test(as.numeric(tmp_results$numSplitBlocks), tmp_results$Fitness_500)
```

# Effect of number of stop codons
```{r}
means = c()
maxs = c()
mins = c()
steps = c()
stops = c()
for (numStop in unique(results$numStop)) {
  tmp = results[results$numStop==numStop,]
  for (i in seq(0,990,10)) {
    m = mean(tmp[,(i/10)+11])
    means = c(means, m)
    sd = sd(tmp[,(i/10)+11])
    mins = c(mins, m-sd)
    maxs = c(maxs, m+sd)
    steps = c(steps, i)
  }
  stops = c(stops, rep(numStop, 100))
}
data = data.frame(numSteps = steps,
                  mean = means,
                  min = mins,
                  max = maxs,
                  numStop = stops)
data$numStop = as.factor(data$numStop)

ggplot(data, aes(x=numSteps, y=mean, color=numStop, fill=numStop)) + 
  geom_line(size=0.4) + 
  scale_color_viridis_d() + scale_fill_viridis_d() +
  theme_test()
```

```{r}
for (nsplit in unique(tmp_results$numSplitBlocks)) {
  print(nsplit)
  tmp_tmp_results = tmp_results[tmp_results$numSplitBlocks==nsplit,]
  print(cor.test(as.numeric(tmp_tmp_results$numStop), tmp_tmp_results$Fitness_500))
}
```

Repeat for N=10, N=100 and N=1,000,000.

# N=10
# Read the data
```{r}
if (file.exists("output/results_10/results_processed.tsv")) {
  results = read.delim("output/results_10/results_processed.tsv", header=TRUE)
  results$numSplitBlocks = as.factor(results$numSplitBlocks)
  results$numStop = as.factor(results$numStop)
} else {
  results = read.delim("output/results_10/results", header=FALSE)
  colnames(results) = c("Id", "Block1", "Block2", "Block3", "Block4", "X", "Z", "B", "J", "Robustness", paste0("Fitness_", seq(0,990,by=10)))
   # number of split codon blocks
  results$numSplitBlocks = as.factor(apply(results, 1, function(x) sum(x[c("X", "Z", "B", "J")]!="-")))
  # number of stop codons
  results$numStop = as.factor(3 - (results$Block2!="-") + 2*apply(results, 1, function(x) sum(x[c("Block1", "Block3", "Block4")]=="O")))
  
  write.table(results, file="output/results_10/results_processed.tsv", quote=FALSE, sep="\t", row.names=FALSE)
}
```

Restrict to only those codes where the size of the summit is the same as in the SGC and the summit is not split:
```{r}
results$nCodons_W = 1+2*(results$Block1=="W") + 1*(results$Block2=="W") + 2*(results$Block3=="W") + 2*(results$Block4=="W")
results$nCodons_L = 6-2*(results$Block1!="-")+1*(results$Block2=="L") 
results$nCodons_A = 4+2*(results$Block3=="A") + 2*(results$Block4=="A")
results$SummitSize = (results$nCodons_W**2) * results$nCodons_A * results$nCodons_L

results$splitW = rowSums(results[,6:9]=="W" )
results$splitL = rowSums(results[,6:9]=="L")
results$splitA = rowSums(results[,6:9]=="A")

tmp_results = results[results$SummitSize==results$SummitSize[results$Id=="standard"] &
                results$splitA==0 & results$splitL==0 & results$splitW==0,]

cor.test(tmp_results$Fitness_500, tmp_results$Robustness)
```

# Effect of number of split codon blocks
```{r}
cor.test(as.numeric(tmp_results$numSplitBlocks), tmp_results$Fitness_500)
```

# Effect of number of stop codons
```{r}
for (nsplit in unique(tmp_results$numSplitBlocks)) {
  print(nsplit)
  tmp_tmp_results = tmp_results[tmp_results$numSplitBlocks==nsplit,]
  print(cor.test(as.numeric(tmp_tmp_results$numStop), tmp_tmp_results$Fitness_500))
}
```

# N=100
# Read the data
```{r}
if (file.exists("output/results_100/results_processed.tsv")) {
  results = read.delim("output/results_100/results_processed.tsv", header=TRUE)
  results$numSplitBlocks = as.factor(results$numSplitBlocks)
  results$numStop = as.factor(results$numStop)
} else {
  results = read.delim("output/results_100/results", header=FALSE)
  colnames(results) = c("Id", "Block1", "Block2", "Block3", "Block4", "X", "Z", "B", "J", "Robustness", paste0("Fitness_", seq(0,990,by=10)))
   # number of split codon blocks
  results$numSplitBlocks = as.factor(apply(results, 1, function(x) sum(x[c("X", "Z", "B", "J")]!="-")))
  # number of stop codons
  results$numStop = as.factor(3 - (results$Block2!="-") + 2*apply(results, 1, function(x) sum(x[c("Block1", "Block3", "Block4")]=="O")))
  
  write.table(results, file="output/results_100/results_processed.tsv", quote=FALSE, sep="\t", row.names=FALSE)
}
```

Restrict to only those codes where the size of the summit is the same as in the SGC and the summit is not split:
```{r}
results$nCodons_W = 1+2*(results$Block1=="W") + 1*(results$Block2=="W") + 2*(results$Block3=="W") + 2*(results$Block4=="W")
results$nCodons_L = 6-2*(results$Block1!="-")+1*(results$Block2=="L") 
results$nCodons_A = 4+2*(results$Block3=="A") + 2*(results$Block4=="A")
results$SummitSize = (results$nCodons_W**2) * results$nCodons_A * results$nCodons_L

results$splitW = rowSums(results[,6:9]=="W" )
results$splitL = rowSums(results[,6:9]=="L")
results$splitA = rowSums(results[,6:9]=="A")

tmp_results = results[results$SummitSize==results$SummitSize[results$Id=="standard"] &
                results$splitA==0 & results$splitL==0 & results$splitW==0,]

cor.test(tmp_results$Fitness_500, tmp_results$Robustness)
```

# Effect of number of split codon blocks
```{r}
cor.test(as.numeric(tmp_results$numSplitBlocks), tmp_results$Fitness_500)
```

# Effect of number of stop codons
```{r}
for (nsplit in unique(tmp_results$numSplitBlocks)) {
  print(nsplit)
  tmp_tmp_results = tmp_results[tmp_results$numSplitBlocks==nsplit,]
  print(cor.test(as.numeric(tmp_tmp_results$numStop), tmp_tmp_results$Fitness_500))
}
```


# N=1,000,000
# Read the data
```{r}
if (file.exists("output/results_1000000/results_processed.tsv")) {
  results = read.delim("output/results_1000000/results_processed.tsv", header=TRUE)
  results$numSplitBlocks = as.factor(results$numSplitBlocks)
  results$numStop = as.factor(results$numStop)
} else {
  results = read.delim("output/results_1000000/results", header=FALSE)
  colnames(results) = c("Id", "Block1", "Block2", "Block3", "Block4", "X", "Z", "B", "J", "Robustness", paste0("Fitness_", seq(0,990,by=10)))
   # number of split codon blocks
  results$numSplitBlocks = as.factor(apply(results, 1, function(x) sum(x[c("X", "Z", "B", "J")]!="-")))
  # number of stop codons
  results$numStop = as.factor(3 - (results$Block2!="-") + 2*apply(results, 1, function(x) sum(x[c("Block1", "Block3", "Block4")]=="O")))
  
  write.table(results, file="output/results_1000000/results_processed.tsv", quote=FALSE, sep="\t", row.names=FALSE)
}
```

Restrict to only those codes where the size of the summit is the same as in the SGC and the summit is not split:
```{r}
results$nCodons_W = 1+2*(results$Block1=="W") + 1*(results$Block2=="W") + 2*(results$Block3=="W") + 2*(results$Block4=="W")
results$nCodons_L = 6-2*(results$Block1!="-")+1*(results$Block2=="L") 
results$nCodons_A = 4+2*(results$Block3=="A") + 2*(results$Block4=="A")
results$SummitSize = (results$nCodons_W**2) * results$nCodons_A * results$nCodons_L

results$splitW = rowSums(results[,6:9]=="W" )
results$splitL = rowSums(results[,6:9]=="L")
results$splitA = rowSums(results[,6:9]=="A")

tmp_results = results[results$SummitSize==results$SummitSize[results$Id=="standard"] &
                results$splitA==0 & results$splitL==0 & results$splitW==0,]

cor.test(tmp_results$Fitness_500, tmp_results$Robustness)
```

# Effect of number of split codon blocks
```{r}
cor.test(as.numeric(tmp_results$numSplitBlocks), tmp_results$Fitness_500)
```

# Effect of number of stop codons
```{r}
for (nsplit in unique(tmp_results$numSplitBlocks)) {
  print(nsplit)
  tmp_tmp_results = tmp_results[tmp_results$numSplitBlocks==nsplit,]
  print(cor.test(as.numeric(tmp_tmp_results$numStop), tmp_tmp_results$Fitness_500))
}
```