---
title: "ParE3, Ostrov codes, landscape ruggedness"
author: "Hana Rozhonova"
date: "6/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(stringr)
library(ggplot2)
library(ggpubr)
library(reshape2)
library(gridExtra)
```

# Read the data
```{r}
results = read.delim("output/results")
colnames(results) = c("Id", "Block1", "Block2", "Block3", "Block4", "X", "Z", "B", "J", "Robustness", "NumPeaks", "MeanPeakHeight", "NoEpi", "MagEpi", "SsEpi", "RsEpi", "PropAccPaths")
# number of split codon blocks
results$numSplitBlocks = apply(results, 1, function(x) sum(x[c("X", "Z", "B", "J")]!="-"))
# number of stop codons
results$numStop = 3 - (results$Block2!="-") + 2*apply(results, 1, function(x) sum(x[c("Block1", "Block3", "Block4")]=="O"))
```

# Correlation of robustness with number of split codon blocks and number of stop codons
```{r}
cor.test(results$Robustness, results$numSplitBlocks)
cor.test(results$Robustness, results$numStop)
```

# Peaks
#### Histogram of the number of peaks
```{r}
peaks_standard = results$NumPeaks[results$Id=="standard"]
p = sum(results$NumPeaks<=peaks_standard)/nrow(results)
ggplot(results, aes(x=NumPeaks)) + 
  geom_histogram(alpha=0.2, color="black") +
  theme_test() +
  geom_vline(xintercept=peaks_standard, linetype="dashed", color="red") +
  labs(x="Number of peaks", y="Count", title=paste0("p=", signif(p, digits=3)))
```

#### Correlation of number of peaks with code robustness
```{r}
(corTest = cor.test(results$NumPeaks, results$Robustness))
```

Density plot:
```{r}
rob_standard = results$Robustness[results$Id=="standard"]
(p_peaks_rob = ggplot(results, aes(x=Robustness, y=NumPeaks)) +
  geom_bin2d(bins=c(22,40), alpha=0.8) +
  scale_fill_continuous(type = "viridis", name="Count") +
  theme_test() +
  theme(legend.position="bottom") +
  geom_point(x=rob_standard, y=peaks_standard, fill="gray", pch=21, size=2) + 
  labs(x="Code robustness", y="Number of local peaks") +
  geom_smooth(color="black", method="lm") +
  geom_text(label=paste0("R = ", signif(corTest$estimate, digits=3)), x=0.39, y=max(results$NumPeaks)-0.03*(max(results$NumPeaks)-min(results$NumPeaks)), size=4) +
  annotate(geom = "label", x=rob_standard, y=peaks_standard-0.07*(max(results$NumPeaks)-min(results$NumPeaks)), label="SGC", alpha=0.8, size=3))
```

#### Number of split codon blocks
Correlation:
```{r}
(corTest = cor.test(results$numSplitBlocks, results$NumPeaks))
```

Boxplot - number of peaks:
```{r}
comparisons_numSplit = list(c(0,1), c(1,2), c(2,3), c(3,4))
ggplot(results, aes(x=numSplitBlocks, y=NumPeaks, group=numSplitBlocks)) + 
  geom_boxplot() +
  geom_point(x=2, y=peaks_standard, color="red") +
  theme_test() +
  labs(x="Number of split codon blocks", y="Number of peaks") +
  stat_compare_means(comparisons = comparisons_numSplit, label = "p.signif")
```


Hold robustness constant, is there still correlation between number of split codon blocks and number of peaks?
```{r}
corrs = c()
pvals = c()
robs = unique(results$Robustness)
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  test = cor.test(tmp$numSplitBlocks, tmp$NumPeaks)
  corrs = c(corrs, test$estimate)
  pvals = c(pvals, test$p.value)
}
data = data.frame(Robustness = robs,
                  Correlation = corrs,
                  Pvalue = pvals)
data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

ggplot(data, aes(x=Robustness, y=Correlation, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of split codon blocks\nand number of peaks", x="Code robustness") +
  scale_shape_discrete(name="P-value")
```

#### Number of stop codons
```{r}
cor.test(results$numStop, results$NumPeaks)
for (numSplit in unique(results$numSplitBlocks)) {
  print(numSplit)
  tmp = results[results$numSplitBlocks==numSplit,]
  print(cor.test(tmp$numStop, tmp$NumPeaks))
}
```

Boxplot - number of peaks (for given number of split codon blocks):
```{r}
for (numSplit in unique(results$numSplitBlocks)) {
  tmp = results[results$numSplitBlocks==numSplit,]
  p= ggplot(tmp, aes(x=numStop, y=NumPeaks, group=numStop)) + 
    geom_boxplot() +
    theme_test() +
    labs(x="Number of stop codons", y="Number of peaks", title=paste(numSplit, "split codon blocks")) +
    stat_compare_means()
  if(numSplit==1) {
    p = p+geom_point(x=2, y=peaks_standard, color="red")
  }
  print(p)
}

```



Hold robustness constant, is there still correlation between number of stop codons and number of peaks?
```{r}
corrs = c()
pvals = c()
numSplit = c()
robs = unique(results$Robustness)
robs2 = c()
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  numsplits = unique(tmp$numSplitBlocks)
  for (ns in numsplits) {
    tmp_tmp = tmp[tmp$numSplitBlocks==ns,]
    test = cor.test(tmp_tmp$numStop, tmp_tmp$NumPeaks)
    corrs = c(corrs, test$estimate)
    pvals = c(pvals, test$p.value)
    numSplit = c(numSplit, ns)
    robs2 = c(robs2, rob)
  }
}
data = data.frame(Robustness = robs2,
                  Correlation = corrs,
                  Pvalue = pvals,
                  NumSplit = numSplit)

data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

data$NumSplit = as.factor(data$NumSplit)
data = data[!is.na(data$Correlation),]
ggplot(data, aes(x=Robustness, y=Correlation, color=NumSplit, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of stop codons\nand number of peaks", x="Code robustness", main="ParE3") +
  scale_color_viridis_d() + geom_hline(yintercept=0, linetype="dashed") +
  scale_shape_discrete(name="P-value")
```

# Accessibility of the global peak
#### Histogram 
```{r}
acc_standard = results$PropAccPaths[results$Id=="standard"]
p = sum(results$PropAccPaths<=acc_standard)/nrow(results)
ggplot(results, aes(x=PropAccPaths)) + 
  geom_histogram(alpha=0.2, color="black") +
  theme_test() +
  geom_vline(xintercept=acc_standard, linetype="dashed", color="red") +
  labs(x="Accessibility of the global peak", y="Count", title=paste0("p=", signif(p, digits=3)))
```

#### Correlation of global peak accessibility with code robustness
```{r}
(corTest = cor.test(results$PropAccPaths, results$Robustness))
```

The plot:
```{r}
(p_paths_rob = ggplot(results, aes(x=Robustness, y=PropAccPaths)) +
  geom_bin2d(bins=c(22,40), alpha=0.8) +
  scale_fill_continuous(type = "viridis", name="Count") +
  theme_test() +
   theme(legend.position="bottom") +
  geom_point(x=rob_standard, y=acc_standard, fill="gray", pch=21, size=2) + 
  labs(x="Code robustness", y="Probability that a randomly\nchosen path to the summit\nis accessible") +
  geom_smooth(color="black") +
  geom_text(label=paste0("R = ", signif(corTest$estimate, digits=3)), x=0.39, y=max(results$PropAccPaths)-0.03*(max(results$PropAccPaths)-min(results$PropAccPaths)), size=4) +
  annotate(geom = "label", x=rob_standard, y=acc_standard-0.07*(max(results$PropAccPaths)-min(results$PropAccPaths)), label="SGC", alpha=0.8, size=3))

```

Restrict to only those codes where the size of the summit is the same as in the SGC and the summit is not split:
```{r}
results$nCodons_D = 2 
results$nCodons_W = 1+2*(results$Block1=="W") + 1*(results$Block2=="W") + 2*(results$Block3=="W") + 2*(results$Block4=="W")
results$nCodons_E = 2+2*(results$Block1=="E") + 1*(results$Block2=="E")  
results$SummitSize = results$nCodons_E * results$nCodons_D * results$nCodons_W

results$splitE = rowSums(results[,6:9]=="E" )
results$splitD = rowSums(results[,6:9]=="D")
results$splitW = rowSums(results[,6:9]=="W")

tmp_results = results[results$SummitSize==results$SummitSize[results$Id=="standard"] &
                results$splitE==0 & results$splitD==0 & results$splitW==0,]

cor.test(tmp_results$PropAccPaths, tmp_results$Robustness)
```


#### Number of split codon blocks
Correlation:
```{r}
(corTest = cor.test(tmp_results$numSplitBlocks, tmp_results$PropAccPaths))
```


Hold robustness constant, is there still correlation between number of split codon blocks and mutational accessibility?
```{r}
corrs = c()
pvals = c()
robs = unique(tmp_results$Robustness)
for (rob in robs) {
  tmp = tmp_results[tmp_results$Robustness==rob,]
  test = cor.test(tmp$numSplitBlocks, tmp$PropAccPaths)
  corrs = c(corrs, test$estimate)
  pvals = c(pvals, test$p.value)
}
data = data.frame(Robustness = robs,
                  Correlation = corrs,
                  Pvalue = pvals)
data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

ggplot(data, aes(x=Robustness, y=Correlation, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of split codon blocks\nand mutational accessibility of the summit", x="Code robustness") +
  scale_shape_discrete(name="P-value") +
  geom_hline(yintercept=0, linetype="dashed")
```

#### Number of stop codons
```{r}
cor.test(tmp_results$numStop, tmp_results$PropAccPaths)
for (numSplit in unique(tmp_results$numSplitBlocks)) {
  print(numSplit)
  tmp = tmp_results[tmp_results$numSplitBlocks==numSplit,]
  print(cor.test(tmp$numStop, tmp$PropAccPaths))
}
```



Hold robustness constant, is there still correlation between number of stop codons and number of peaks?
```{r}
corrs = c()
pvals = c()
numSplit = c()
robs = unique(tmp_results$Robustness)
robs2 = c()
for (rob in robs) {
  tmp = tmp_results[tmp_results$Robustness==rob,]
  numsplits = unique(tmp$numSplitBlocks)
  for (ns in numsplits) {
    tmp_tmp = tmp[tmp$numSplitBlocks==ns,]
    if (nrow(tmp_tmp)>2) {
      test = cor.test(tmp_tmp$numStop, tmp_tmp$PropAccPaths)
      corrs = c(corrs, test$estimate)
      pvals = c(pvals, test$p.value)
      numSplit = c(numSplit, ns)
      robs2 = c(robs2, rob)
    }
  }
}
data = data.frame(Robustness = robs2,
                  Correlation = corrs,
                  Pvalue = pvals,
                  NumSplit = numSplit)

data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

data$NumSplit = as.factor(data$NumSplit)
data = data[!is.na(data$Correlation),]
ggplot(data, aes(x=Robustness, y=Correlation, color=NumSplit, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of stop codons\nand mutational accessibility of the summit", x="Code robustness", main="ParE3") +
  scale_color_viridis_d() + geom_hline(yintercept=0, linetype="dashed") +
  scale_shape_discrete(name="P-value")
```

# Epistasis
## No epistasis
#### Correlation with code robustness
```{r}
(corTest = cor.test(results$NoEpi, results$Robustness))
```


#### Number of split codon blocks
Correlation:
```{r}
(corTest = cor.test(results$numSplitBlocks, results$NoEpi))
```

Hold robustness constant, is there still correlation between number of split codon blocks and no epistasis?
```{r}
corrs = c()
pvals = c()
robs = unique(results$Robustness)
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  test = cor.test(tmp$numSplitBlocks, tmp$NoEpi)
  corrs = c(corrs, test$estimate)
  pvals = c(pvals, test$p.value)
}
data = data.frame(Robustness = robs,
                  Correlation = corrs,
                  Pvalue = pvals)

data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

ggplot(data, aes(x=Robustness, y=Correlation, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of split codon blocks\nand prevalence of no epistasis", x="Code robustness") +
  scale_shape_discrete(name="P-value")
```

#### Number of stop codons
```{r}
cor.test(results$numStop, results$NoEpi)
for (numSplit in unique(results$numSplitBlocks)) {
  print(numSplit)
  tmp = results[results$numSplitBlocks==numSplit,]
  print(cor.test(tmp$numStop, tmp$NoEpi))
}
```



Hold robustness constant, is there still correlation?
```{r}
corrs = c()
pvals = c()
numSplit = c()
robs = unique(results$Robustness)
robs2 = c()
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  numsplits = unique(tmp$numSplitBlocks)
  for (ns in numsplits) {
    tmp_tmp = tmp[tmp$numSplitBlocks==ns,]
    test = cor.test(tmp_tmp$numStop, tmp_tmp$NoEpi)
    corrs = c(corrs, test$estimate)
    pvals = c(pvals, test$p.value)
    numSplit = c(numSplit, ns)
    robs2 = c(robs2, rob)
  }
}
data = data.frame(Robustness = robs2,
                  Correlation = corrs,
                  Pvalue = pvals,
                  NumSplit = numSplit)

data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

data$NumSplit = as.factor(data$NumSplit)
data = data[!is.na(data$Correlation),]
 ggplot(data, aes(x=Robustness, y=Correlation, color=NumSplit, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of stop codons\nand prevalence of no epistasis", x="Code robustness", main="ParE3") +
  scale_color_viridis_d() + geom_hline(yintercept=0, linetype="dashed") +
  scale_shape_discrete(name="P-value")
```

## Magnitude epistasis
#### Correlation with code robustness
```{r}
(corTest = cor.test(results$MagEpi, results$Robustness))
```


#### Number of split codon blocks
Correlation:
```{r}
(corTest = cor.test(results$numSplitBlocks, results$MagEpi))
```

Hold robustness constant, is there still correlation between number of split codon blocks and no epistasis?
```{r}
corrs = c()
pvals = c()
robs = unique(results$Robustness)
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  test = cor.test(tmp$numSplitBlocks, tmp$MagEpi)
  corrs = c(corrs, test$estimate)
  pvals = c(pvals, test$p.value)
}
data = data.frame(Robustness = robs,
                  Correlation = corrs,
                  Pvalue = pvals)

data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

ggplot(data, aes(x=Robustness, y=Correlation, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of split codon blocks\nand prevalence of magnitude epistasis", x="Code robustness") +
  scale_shape_discrete(name="P-value")
```

#### Number of stop codons
```{r}
cor.test(results$numStop, results$MagEpi)
for (numSplit in unique(results$numSplitBlocks)) {
  print(numSplit)
  tmp = results[results$numSplitBlocks==numSplit,]
  print(cor.test(tmp$numStop, tmp$MagEpi))
}
```



Hold robustness constant, is there still correlation?
```{r}
corrs = c()
pvals = c()
numSplit = c()
robs = unique(results$Robustness)
robs2 = c()
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  numsplits = unique(tmp$numSplitBlocks)
  for (ns in numsplits) {
    tmp_tmp = tmp[tmp$numSplitBlocks==ns,]
    test = cor.test(tmp_tmp$numStop, tmp_tmp$MagEpi)
    corrs = c(corrs, test$estimate)
    pvals = c(pvals, test$p.value)
    numSplit = c(numSplit, ns)
    robs2 = c(robs2, rob)
  }
}
data = data.frame(Robustness = robs2,
                  Correlation = corrs,
                  Pvalue = pvals,
                  NumSplit = numSplit)

data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

data$NumSplit = as.factor(data$NumSplit)
data = data[!is.na(data$Correlation),]
ggplot(data, aes(x=Robustness, y=Correlation, color=NumSplit, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of stop codons\nand prevalence of magnitude epistasis", x="Code robustness", main="ParE3") +
  scale_color_viridis_d() + geom_hline(yintercept=0, linetype="dashed") +
  scale_shape_discrete(name="P-value")
```

## Simple-sign epistasis
#### Correlation with code robustness
```{r}
(corTest = cor.test(results$SsEpi, results$Robustness))
```


#### Number of split codon blocks
Correlation:
```{r}
(corTest = cor.test(results$numSplitBlocks, results$SsEpi))
```

Hold robustness constant, is there still correlation between number of split codon blocks and no epistasis?
```{r}
corrs = c()
pvals = c()
robs = unique(results$Robustness)
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  test = cor.test(tmp$numSplitBlocks, tmp$SsEpi)
  corrs = c(corrs, test$estimate)
  pvals = c(pvals, test$p.value)
}
data = data.frame(Robustness = robs,
                  Correlation = corrs,
                  Pvalue = pvals)

data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

ggplot(data, aes(x=Robustness, y=Correlation, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of split codon blocks\nand prevalence of simple-sign epistasis", x="Code robustness") +
  scale_shape_discrete(name="P-value")
```

#### Number of stop codons
```{r}
cor.test(results$numStop, results$SsEpi)
for (numSplit in unique(results$numSplitBlocks)) {
  print(numSplit)
  tmp = results[results$numSplitBlocks==numSplit,]
  print(cor.test(tmp$numStop, tmp$SsEpi))
}
```



Hold robustness constant, is there still correlation?
```{r}
corrs = c()
pvals = c()
numSplit = c()
robs = unique(results$Robustness)
robs2 = c()
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  numsplits = unique(tmp$numSplitBlocks)
  for (ns in numsplits) {
    tmp_tmp = tmp[tmp$numSplitBlocks==ns,]
    test = cor.test(tmp_tmp$numStop, tmp_tmp$SsEpi)
    corrs = c(corrs, test$estimate)
    pvals = c(pvals, test$p.value)
    numSplit = c(numSplit, ns)
    robs2 = c(robs2, rob)
  }
}
data = data.frame(Robustness = robs2,
                  Correlation = corrs,
                  Pvalue = pvals,
                  NumSplit = numSplit)

data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

data$NumSplit = as.factor(data$NumSplit)
data = data[!is.na(data$Correlation),]
ggplot(data, aes(x=Robustness, y=Correlation, color=NumSplit, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of stop codons\nand prevalence of simple-sign epistasis", x="Code robustness", main="ParE3") +
  scale_color_viridis_d() + geom_hline(yintercept=0, linetype="dashed") +
  scale_shape_discrete(name="P-value")
```

## Reciprocal-sign epistasis
#### Correlation with code robustness
```{r}
(corTest = cor.test(results$RsEpi, results$Robustness))
```





#### Number of split codon blocks
Correlation:
```{r}
(corTest = cor.test(results$numSplitBlocks, results$RsEpi))
```

Hold robustness constant, is there still correlation between number of split codon blocks and no epistasis?
```{r}
corrs = c()
pvals = c()
robs = unique(results$Robustness)
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  test = cor.test(tmp$numSplitBlocks, tmp$RsEpi)
  corrs = c(corrs, test$estimate)
  pvals = c(pvals, test$p.value)
}
data = data.frame(Robustness = robs,
                  Correlation = corrs,
                  Pvalue = pvals)

data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

ggplot(data, aes(x=Robustness, y=Correlation, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of split codon blocks\nand prevalence of reciprocal-sign epistasis", x="Code robustness") +
  scale_shape_discrete(name="P-value")
```

#### Number of stop codons
```{r}
cor.test(results$numStop, results$RsEpi)
for (numSplit in unique(results$numSplitBlocks)) {
  print(numSplit)
  tmp = results[results$numSplitBlocks==numSplit,]
  print(cor.test(tmp$numStop, tmp$RsEpi))
}
```



Hold robustness constant, is there still correlation?
```{r}
corrs = c()
pvals = c()
numSplit = c()
robs = unique(results$Robustness)
robs2 = c()
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  numsplits = unique(tmp$numSplitBlocks)
  for (ns in numsplits) {
    tmp_tmp = tmp[tmp$numSplitBlocks==ns,]
    test = cor.test(tmp_tmp$numStop, tmp_tmp$RsEpi)
    corrs = c(corrs, test$estimate)
    pvals = c(pvals, test$p.value)
    numSplit = c(numSplit, ns)
    robs2 = c(robs2, rob)
  }
}
data = data.frame(Robustness = robs2,
                  Correlation = corrs,
                  Pvalue = pvals,
                  NumSplit = numSplit)

data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

data$NumSplit = as.factor(data$NumSplit)
data = data[!is.na(data$Correlation),]
ggplot(data, aes(x=Robustness, y=Correlation, color=NumSplit, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of stop codons\nand prevalence of reciprocal-sign epistasis", x="Code robustness", main="ParE3") +
  scale_color_viridis_d() + geom_hline(yintercept=0, linetype="dashed") +
  scale_shape_discrete(name="P-value")
```


## Normalize by the number of epistatic squares
```{r}
results$MagEpi = results$MagEpi/(1-results$NoEpi)
results$SsEpi = results$SsEpi/(1-results$NoEpi)
results$RsEpi = results$RsEpi/(1-results$NoEpi)
```


## Magnitude epistasis
#### Correlation with code robustness
```{r}
(corTest = cor.test(results$MagEpi, results$Robustness))
```


#### Number of split codon blocks
Correlation:
```{r}
(corTest = cor.test(results$numSplitBlocks, results$MagEpi))
```

Hold robustness constant, is there still correlation between number of split codon blocks and no epistasis?
```{r}
corrs = c()
pvals = c()
robs = unique(results$Robustness)
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  test = cor.test(tmp$numSplitBlocks, tmp$MagEpi)
  corrs = c(corrs, test$estimate)
  pvals = c(pvals, test$p.value)
}
data = data.frame(Robustness = robs,
                  Correlation = corrs,
                  Pvalue = pvals)

data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

ggplot(data, aes(x=Robustness, y=Correlation, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of split codon blocks\nand prevalence of magnitude epistasis", x="Code robustness") +
  scale_shape_discrete(name="P-value")

```

#### Number of stop codons
```{r}
cor.test(results$numStop, results$MagEpi)
for (numSplit in unique(results$numSplitBlocks)) {
  print(numSplit)
  tmp = results[results$numSplitBlocks==numSplit,]
  print(cor.test(tmp$numStop, tmp$MagEpi))
}
```



Hold robustness constant, is there still correlation?
```{r}
corrs = c()
pvals = c()
numSplit = c()
robs = unique(results$Robustness)
robs2 = c()
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  numsplits = unique(tmp$numSplitBlocks)
  for (ns in numsplits) {
    tmp_tmp = tmp[tmp$numSplitBlocks==ns,]
    test = cor.test(tmp_tmp$numStop, tmp_tmp$MagEpi)
    corrs = c(corrs, test$estimate)
    pvals = c(pvals, test$p.value)
    numSplit = c(numSplit, ns)
    robs2 = c(robs2, rob)
  }
}
data = data.frame(Robustness = robs2,
                  Correlation = corrs,
                  Pvalue = pvals,
                  NumSplit = numSplit)

data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

data$NumSplit = as.factor(data$NumSplit)
data = data[!is.na(data$Correlation),]
ggplot(data, aes(x=Robustness, y=Correlation, color=NumSplit, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of stop codons\nand prevalence of magnitude epistasis", x="Code robustness", main="ParE3") +
  scale_color_viridis_d() + geom_hline(yintercept=0, linetype="dashed") +
  scale_shape_discrete(name="P-value")
```

## Simple-sign epistasis
#### Correlation with code robustness
```{r}
(corTest = cor.test(results$SsEpi, results$Robustness))
```


#### Number of split codon blocks
Correlation:
```{r}
(corTest = cor.test(results$numSplitBlocks, results$SsEpi))
```

Hold robustness constant, is there still correlation between number of split codon blocks and no epistasis?
```{r}
corrs = c()
pvals = c()
robs = unique(results$Robustness)
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  test = cor.test(tmp$numSplitBlocks, tmp$SsEpi)
  corrs = c(corrs, test$estimate)
  pvals = c(pvals, test$p.value)
}
data = data.frame(Robustness = robs,
                  Correlation = corrs,
                  Pvalue = pvals)

data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

ggplot(data, aes(x=Robustness, y=Correlation, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of split codon blocks\nand prevalence of simple-sign epistasis", x="Code robustness") +
  scale_shape_discrete(name="P-value") +
  geom_hline(yintercept=0, linetype="dashed")
```

#### Number of stop codons
```{r}
cor.test(results$numStop, results$SsEpi)
for (numSplit in unique(results$numSplitBlocks)) {
  print(numSplit)
  tmp = results[results$numSplitBlocks==numSplit,]
  print(cor.test(tmp$numStop, tmp$SsEpi))
}
```



Hold robustness constant, is there still correlation?
```{r}
corrs = c()
pvals = c()
numSplit = c()
robs = unique(results$Robustness)
robs2 = c()
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  numsplits = unique(tmp$numSplitBlocks)
  for (ns in numsplits) {
    tmp_tmp = tmp[tmp$numSplitBlocks==ns,]
    test = cor.test(tmp_tmp$numStop, tmp_tmp$SsEpi)
    corrs = c(corrs, test$estimate)
    pvals = c(pvals, test$p.value)
    numSplit = c(numSplit, ns)
    robs2 = c(robs2, rob)
  }
}
data = data.frame(Robustness = robs2,
                  Correlation = corrs,
                  Pvalue = pvals,
                  NumSplit = numSplit)

data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

data$NumSplit = as.factor(data$NumSplit)
data = data[!is.na(data$Correlation),]
ggplot(data, aes(x=Robustness, y=Correlation, color=NumSplit, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of stop codons\nand prevalence of simple-sign epistasis", x="Code robustness", main="ParE3") +
  scale_color_viridis_d() + geom_hline(yintercept=0, linetype="dashed") +
  scale_shape_discrete(name="P-value")
```

## Reciprocal-sign epistasis
#### Correlation with code robustness
```{r}
(corTest = cor.test(results$RsEpi, results$Robustness))
```

```{r}
rsEpi_standard = results$RsEpi[results$Id=="standard"]
(p_rsEpi_rob = ggplot(results, aes(x=Robustness, y=RsEpi)) +
  geom_bin2d(bins=c(22,40), alpha=0.8) +
  scale_fill_continuous(type = "viridis", name="Count") +
  theme_test() +
  theme(legend.position="bottom") +
  geom_point(x=rob_standard, y=rsEpi_standard, fill="gray", pch=21, size=2) + 
  labs(x="Code robustness", y="Proportion of reciprocal-sign\nepistasis") +
  geom_smooth(color="black") +
  #theme(legend.position="none") +
  geom_text(label=paste0("R = ", signif(corTest$estimate, digits=3)), x=0.39, y=max(results$RsEpi)-0.03*(max(results$RsEpi)-min(results$RsEpi)), size=4) +
  annotate(geom = "label", x=rob_standard, y=rsEpi_standard-0.07*(max(results$RsEpi)-min(results$RsEpi)), label="SGC", alpha=0.8, size=3))
```


#### Number of split codon blocks
Correlation:
```{r}
(corTest = cor.test(results$numSplitBlocks, results$RsEpi))
```

Hold robustness constant, is there still correlation between number of split codon blocks and no epistasis?
```{r}
corrs = c()
pvals = c()
robs = unique(results$Robustness)
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  test = cor.test(tmp$numSplitBlocks, tmp$RsEpi)
  corrs = c(corrs, test$estimate)
  pvals = c(pvals, test$p.value)
}
data = data.frame(Robustness = robs,
                  Correlation = corrs,
                  Pvalue = pvals)

data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

ggplot(data, aes(x=Robustness, y=Correlation, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of split codon blocks\nand prevalence of reciprocal-sign epistasis", x="Code robustness") +
  scale_shape_discrete(name="P-value")
```

#### Number of stop codons
```{r}
cor.test(results$numStop, results$RsEpi)
for (numSplit in unique(results$numSplitBlocks)) {
  print(numSplit)
  tmp = results[results$numSplitBlocks==numSplit,]
  print(cor.test(tmp$numStop, tmp$RsEpi))
}
```



Hold robustness constant, is there still correlation?
```{r}
corrs = c()
pvals = c()
numSplit = c()
robs = unique(results$Robustness)
robs2 = c()
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  numsplits = unique(tmp$numSplitBlocks)
  for (ns in numsplits) {
    tmp_tmp = tmp[tmp$numSplitBlocks==ns,]
    test = cor.test(tmp_tmp$numStop, tmp_tmp$RsEpi)
    corrs = c(corrs, test$estimate)
    pvals = c(pvals, test$p.value)
    numSplit = c(numSplit, ns)
    robs2 = c(robs2, rob)
  }
}
data = data.frame(Robustness = robs2,
                  Correlation = corrs,
                  Pvalue = pvals,
                  NumSplit = numSplit)

data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

data$NumSplit = as.factor(data$NumSplit)
data = data[!is.na(data$Correlation),]
ggplot(data, aes(x=Robustness, y=Correlation, color=NumSplit, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of stop codons\nand prevalence of reciprocal-sign epistasis", x="Code robustness", main="ParE3") +
  scale_color_viridis_d() + geom_hline(yintercept=0, linetype="dashed") +
  scale_shape_discrete(name="P-value")
```

