---
title: 'ParE2: Ostrov codes, greedy walks'
author: "Hana Rozhonova"
date: "6/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(stringr)
library(ggplot2)
library(gridExtra)
library(ggpubr)
```

# Read the data
```{r}
if (file.exists("output/results_processed.tsv")) {
  results = read.delim("output/results_processed.tsv", header=TRUE)
} else {
  results = read.delim("output/results", header=FALSE, stringsAsFactors = FALSE)
  colnames(results) = c("Id", "Block1", "Block2", "Block3", "Block4", "X", "Z", "B", "J", "Robustness", "MeanFitness", "MeanSteps", "Peaks")
  
  # number of split codon blocks
  results$numSplitBlocks = apply(results, 1, function(x) sum(x[c("X", "Z", "B", "J")]!="-"))
  # number of stop codons
  results$numStop = 3 - (results$Block2!="-") + 2*apply(results, 1, function(x) sum(x[c("Block1", "Block3", "Block4")]=="O"))
  
  # entropy of the reached peaks
  entropy_peaks = c()
  for (i in 1:nrow(results)) {
    peaks_counts = str_split(results$Peaks[i], ",")[[1]]
    peaks_counts = head(peaks_counts, -1)
    peaks_counts = as.numeric(lapply(str_split(peaks_counts, ":"), function(x) x[2]))
    # compute Shannon entropy
    peaks_probs = peaks_counts/sum(peaks_counts)
    entropy = -1*sum(peaks_probs*log(peaks_probs))
    entropy_peaks = c(entropy_peaks, entropy)
  }
  results$Entropy_peaks = entropy_peaks
  
  # probability of reaching the global peak
  prop = function(peaks, opt) {
    peaks_split = strsplit(peaks, ",")[[1]]
    peak_opt = peaks_split[startsWith(peaks_split, opt)]
    peak_opt_count = sum(as.numeric(unlist(strsplit(peak_opt, ":"))[seq(2, 2*length(peak_opt), 2)]))
    count_total = 0
    for (peak in peaks_split) {
      count_total = count_total + as.numeric(strsplit(peak, ":")[[1]][2])
    }
    return(peak_opt_count/count_total)
  }
  # global max
  data = read.delim("input/map.tsv", header=TRUE)
  global_max = as.character(data$sequence[which.max(data$phenotype)])
  results$Prob_reaching_global_peak = sapply(results$Peaks, prop, global_max)
  write.table(results, file="output/results_processed.tsv", quote=FALSE, sep="\t", row.names=FALSE)
}
```

# Histograms: Mean fitness, mean number of steps, entropy of the reached peaks, and probability of reaching the global peak(s)
```{r}
fit_standard = results$MeanFitness[results$Id=="standard"]
p = sum(results$MeanFitness>=fit_standard)/nrow(results)
hist_fit = ggplot(results, aes(x=MeanFitness)) + 
  geom_histogram(alpha=0.2, color="black") +
  theme_test() +
  geom_vline(xintercept=fit_standard, linetype="dashed", color="red") +
  labs(x="Mean fitness", y="Count", title=paste0("p=", signif(p, digits=3)))

steps_standard = results$MeanSteps[results$Id=="standard"]
p = sum(results$MeanSteps>=steps_standard)/nrow(results)
hist_steps = ggplot(results, aes(x=MeanSteps)) + 
  geom_histogram(alpha=0.2, color="black") +
  theme_test() +
  geom_vline(xintercept=steps_standard, linetype="dashed", color="red") +
  labs(x="Mean number of steps", y="Count", title=paste0("p=", signif(p, digits=3)))

ent_standard = results$Entropy_peaks[results$Id=="standard"]
p = sum(results$Entropy_peaks<=ent_standard)/nrow(results)
hist_ent = ggplot(results, aes(x=Entropy_peaks)) + 
  geom_histogram(alpha=0.2, color="black") +
  theme_test() +
  geom_vline(xintercept=ent_standard, linetype="dashed", color="red") +
  labs(x="Entropy of the distribution of reached peaks", y="Count", title=paste0("p=", signif(p, digits=3)))

prob_standard = results$Prob_reaching_global_peak[results$Id=="standard"]
p = sum(results$Prob_reaching_global_peak>=prob_standard)/nrow(results)
hist_prob = ggplot(results, aes(x=Prob_reaching_global_peak)) + 
  geom_histogram(alpha=0.2, color="black") +
  theme_test() +
  geom_vline(xintercept=prob_standard, linetype="dashed", color="red") +
  labs(x="Probability of reaching the global peak(s)", y="Count", title=paste0("p=", signif(p, digits=3)))

grid.arrange(hist_fit, hist_steps, hist_ent, hist_prob, ncol=2)
```

# Correlations with code robustness
```{r}
rob_standard = results$Robustness[results$Id=="standard"]

(corTest = cor.test(results$MeanFitness, results$Robustness))
p_fit = ggplot(results, aes(x=Robustness, y=MeanFitness)) +
  geom_bin2d(bins=30) +
  scale_fill_continuous(type = "viridis") +
  theme_test() +
  geom_point(x=rob_standard, y=fit_standard, color="red") + 
  labs(x="Code robustness", y="Mean fitness", title=paste0("R=", signif(corTest$estimate, digits=3)))

(corTest = cor.test(results$MeanSteps, results$Robustness))
p_steps = ggplot(results, aes(x=Robustness, y=MeanSteps)) +
  geom_bin2d(bins=30) +
  scale_fill_continuous(type = "viridis") +
  theme_test() +
  geom_point(x=rob_standard, y=steps_standard, color="red") + 
  labs(x="Code robustness", y="Mean number of steps", title=paste0("R=", signif(corTest$estimate, digits=3)))

(corTest = cor.test(results$Entropy_peaks, results$Robustness))
p_ent = ggplot(results, aes(x=Robustness, y=Entropy_peaks)) +
  geom_bin2d(bins=30) +
  scale_fill_continuous(type = "viridis") +
  theme_test() +
  geom_point(x=rob_standard, y=ent_standard, color="red") + 
  labs(x="Code robustness", y="Entropy of the distribution of reached peaks", title=paste0("R=", signif(corTest$estimate, digits=3)))

(corTest = cor.test(results$Prob_reaching_global_peak, results$Robustness))
p_prob = ggplot(results, aes(x=Robustness, y=Prob_reaching_global_peak)) +
  geom_bin2d(bins=30) +
  scale_fill_continuous(type = "viridis") +
  theme_test() +
  geom_point(x=rob_standard, y=prob_standard, color="red") + 
  labs(x="Code robustness", y="Probability of reaching the global peak", title=paste0("R=", signif(corTest$estimate, digits=3)))

grid.arrange(p_fit, p_steps, p_ent, p_prob, nrow=2)
```


Restrict to only those codes where the size of the summit is the same as in the SGC and the summit is not split:
```{r}
results$nCodons_E = 2+2*(results$Block1=="E") + 1*(results$Block2=="E") 
results$nCodons_L = 6-2*(results$Block1!="-")+1*(results$Block2=="L") 
results$nCodons_K = 2+2*(results$Block1=="K") + 1*(results$Block2=="K") 
results$SummitSize = results$nCodons_E * results$nCodons_L * results$nCodons_K

results$splitE = rowSums(results[,6:9]=="E" )
results$splitL = rowSums(results[,6:9]=="L")
results$splitK = rowSums(results[,6:9]=="K")

tmp_results = results[results$SummitSize==results$SummitSize[results$Id=="standard"] &
                results$splitE==0 & results$splitL==0 & results$splitK==0,]

cor.test(tmp_results$MeanFitness, tmp_results$Robustness)
cor.test(tmp_results$MeanSteps, tmp_results$Robustness)
cor.test(tmp_results$Entropy_peaks, tmp_results$Robustness)
```

# Correlation with number of split codon blocks
```{r}
cor.test(results$numSplitBlocks, results$MeanFitness)
comparisons_numSplit = list(c(0, 1), c(1, 2), c(2,3), c(3,4))
p_fit = ggplot(results, aes(x=numSplitBlocks, y=MeanFitness, group=numSplitBlocks)) + 
  geom_boxplot() +
  geom_point(x=2, y=fit_standard, color="red") +
  theme_test() +
  labs(x="Number of split codon blocks", y="Mean fitness") +
  stat_compare_means(comparisons = comparisons_numSplit, label = "p.signif")

cor.test(results$numSplitBlocks, results$MeanSteps)
p_steps = ggplot(results, aes(x=numSplitBlocks, y=MeanSteps,  group=numSplitBlocks)) + 
  geom_boxplot() +
  geom_point(x=2, y=steps_standard, color="red") +
  theme_test() +
  labs(x="Number of split codon blocks", y="Mean number of steps") +
  stat_compare_means(comparisons = comparisons_numSplit, label = "p.signif")

cor.test(results$numSplitBlocks, results$Entropy_peaks)
p_ent = ggplot(results, aes(x=numSplitBlocks, y=Entropy_peaks,  group=numSplitBlocks)) + 
  geom_boxplot() +
  geom_point(x=2, y=ent_standard, color="red") +
  theme_test() +
  labs(x="Number of split codon blocks", y="Entropy of the distribution of reached peaks") +
  stat_compare_means(comparisons = comparisons_numSplit, label = "p.signif")

p_prob = ggplot(results, aes(x=numSplitBlocks, y=Prob_reaching_global_peak,  group=numSplitBlocks)) + 
  geom_boxplot() +
  geom_point(x=2, y=prob_standard, color="red") +
  theme_test() +
  labs(x="Number of split codon blocks", y="Probability of reaching the global peak") +
  stat_compare_means(comparisons = comparisons_numSplit, label = "p.signif")

grid.arrange(p_fit, p_steps, p_ent, p_prob, nrow=2)
```

Hold robustness constant, is there still correlation?
```{r}
# mean fitness
corrs = c()
pvals = c()
robs = unique(results$Robustness)
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  test = cor.test(tmp$numSplitBlocks, tmp$MeanFitness)
  corrs = c(corrs, test$estimate)
  pvals = c(pvals, test$p.value)
}
data = data.frame(Robustness = robs,
                  Correlation = corrs,
                  Pvalue = pvals)
data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

ggplot(data, aes(x=Robustness, y=Correlation, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of split codon blocks\nand mean fitness", x="Code robustness") +
  scale_shape_discrete(name="P-value") +
  geom_hline(yintercept=0, linetype="dashed")

# mean number of steps
corrs = c()
pvals = c()
robs = unique(results$Robustness)
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  test = cor.test(tmp$numSplitBlocks, tmp$MeanSteps)
  corrs = c(corrs, test$estimate)
  pvals = c(pvals, test$p.value)
}
data = data.frame(Robustness = robs,
                  Correlation = corrs,
                  Pvalue = pvals)
data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

ggplot(data, aes(x=Robustness, y=Correlation, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of split codon blocks\nand mean number of steps", x="Code robustness") +
  scale_shape_discrete(name="P-value") +
  geom_hline(yintercept=0, linetype="dashed")

# entropy
corrs = c()
pvals = c()
robs = unique(results$Robustness)
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  test = cor.test(tmp$numSplitBlocks, tmp$Entropy_peaks)
  corrs = c(corrs, test$estimate)
  pvals = c(pvals, test$p.value)
}
data = data.frame(Robustness = robs,
                  Correlation = corrs,
                  Pvalue = pvals)
data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

ggplot(data, aes(x=Robustness, y=Correlation, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of split codon blocks\nand entropy of the distribution of reached peaks", x="Code robustness") +
  scale_shape_discrete(name="P-value") +
  geom_hline(yintercept=0, linetype="dashed")
```

# Number of stop codons
```{r}
for (numSplit in unique(results$numSplitBlocks)) {
  tmp = results[results$numSplitBlocks==numSplit,]
  print(numSplit)
  print(cor.test(tmp$numStop, tmp$MeanFitness))
}
```

```{r}
for (numSplit in unique(results$numSplitBlocks)) {
  tmp = results[results$numSplitBlocks==numSplit,]
  print(numSplit)
  print(cor.test(tmp$numStop, tmp$MeanSteps))
}
```

```{r}
for (numSplit in unique(results$numSplitBlocks)) {
  tmp = results[results$numSplitBlocks==numSplit,]
  print(numSplit)
  print(cor.test(tmp$numStop, tmp$Entropy_peaks))
}
```

```{r}
for (numSplit in c(0, 1, 2, 3)) {
  tmp = results[results$numSplitBlocks==numSplit,]
  numStops = sort(unique(tmp$numStop))
  comparisons_numStop = list()
  for (i in 2:length(numStops)) {
    comparisons_numStop = append(comparisons_numStop, list(c(numStops[i-1], numStops[i])))
  }
  p_fit = ggplot(tmp, aes(x=numStop, y=MeanFitness,  group=numStop)) + 
    geom_boxplot() +
    theme_test() +
    labs(x="Number of stop codons", y="Mean fitness") +
    stat_compare_means(comparisons = comparisons_numStop, label = "p.signif")
  if (numSplit=="1") p_fit = p_fit + geom_point(x=2, y=fit_standard, color="red")
  
  p_steps = ggplot(tmp, aes(x=numStop, y=MeanSteps,  group=numStop)) + 
    geom_boxplot() +
    theme_test() +
    labs(x="Number of stop codons", y="Mean number of steps") +
    stat_compare_means(comparisons = comparisons_numStop, label = "p.signif")
  if (numSplit=="1") p_steps = p_steps + geom_point(x=2, y=steps_standard, color="red")
  
  p_ent = ggplot(tmp, aes(x=numStop, y=Entropy_peaks,  group=numStop)) + 
    geom_boxplot() +
    theme_test() +
    labs(x="Number of stop codons", y="Entropy of the distribution of reached peaks") +
    stat_compare_means(comparisons = comparisons_numStop, label = "p.signif")
  if (numSplit=="1") p_ent = p_ent + geom_point(x=2, y=ent_standard, color="red") 
  
  p_prob = ggplot(tmp, aes(x=numStop, y=Prob_reaching_global_peak,  group=numStop)) + 
    geom_boxplot() +
    theme_test() +
    labs(x="Number of stop codons", y="Probability of reaching the global peak") +
    stat_compare_means(comparisons = comparisons_numStop, label = "p.signif")
  if (numSplit=="1") p_prob = p_prob + geom_point(x=2, y=prob_standard, color="red") 
  
  print(grid.arrange(p_fit, p_steps, p_ent, p_prob, nrow=2, top=paste(numSplit, "split codon blocks")))
}
```

Hold robustness constant, is there still correlation between number of stop codons and greedy walks results?
```{r}
# mean fitness
corrs = c()
pvals = c()
numSplit = c()
robs = unique(results$Robustness)
robs2 = c()
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  numsplits = unique(tmp$numSplitBlocks)
  for (ns in numsplits) {
    tmp_tmp = tmp[tmp$numSplitBlocks==ns,]
    test = cor.test(tmp_tmp$numStop, tmp_tmp$MeanFitness)
    corrs = c(corrs, test$estimate)
    pvals = c(pvals, test$p.value)
    numSplit = c(numSplit, ns)
    robs2 = c(robs2, rob)
  }
}
data = data.frame(Robustness = robs2,
                  Correlation = corrs,
                  Pvalue = pvals,
                  NumSplit = numSplit)
data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

data$NumSplit = as.factor(data$NumSplit)
data = data[!is.na(data$Correlation),]
ggplot(data, aes(x=Robustness, y=Correlation, color=NumSplit, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of stop codons\nand mean fitness", x="Code robustness") +
  scale_color_viridis_d() + geom_hline(yintercept=0, linetype="dashed") +
  scale_shape_discrete(name="P-value")

# mean number of steps
corrs = c()
pvals = c()
numSplit = c()
robs = unique(results$Robustness)
robs2 = c()
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  numsplits = unique(tmp$numSplitBlocks)
  for (ns in numsplits) {
    tmp_tmp = tmp[tmp$numSplitBlocks==ns,]
    test = cor.test(tmp_tmp$numStop, tmp_tmp$MeanSteps)
    corrs = c(corrs, test$estimate)
    pvals = c(pvals, test$p.value)
    numSplit = c(numSplit, ns)
    robs2 = c(robs2, rob)
  }
}
data = data.frame(Robustness = robs2,
                  Correlation = corrs,
                  Pvalue = pvals,
                  NumSplit = numSplit)
data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

data$NumSplit = as.factor(data$NumSplit)
data = data[!is.na(data$Correlation),]
ggplot(data, aes(x=Robustness, y=Correlation, color=NumSplit, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of stop codons\nand mean number of steps", x="Code robustness") +
  scale_color_viridis_d() + geom_hline(yintercept=0, linetype="dashed") +
  scale_shape_discrete(name="P-value")

# entropy
corrs = c()
pvals = c()
numSplit = c()
robs = unique(results$Robustness)
robs2 = c()
for (rob in robs) {
  tmp = results[results$Robustness==rob,]
  numsplits = unique(tmp$numSplitBlocks)
  for (ns in numsplits) {
    tmp_tmp = tmp[tmp$numSplitBlocks==ns,]
    test = cor.test(tmp_tmp$numStop, tmp_tmp$Entropy_peaks)
    corrs = c(corrs, test$estimate)
    pvals = c(pvals, test$p.value)
    numSplit = c(numSplit, ns)
    robs2 = c(robs2, rob)
  }
}
data = data.frame(Robustness = robs2,
                  Correlation = corrs,
                  Pvalue = pvals,
                  NumSplit = numSplit)
data$pval_cat = ifelse(data$Pvalue>=0.05, ">=0.05",
                       ifelse(data$Pvalue>=0.01, ">=0.01",
                              ifelse(data$Pvalue>=0.001, ">=0.001", "<0.001")))

data$NumSplit = as.factor(data$NumSplit)
data = data[!is.na(data$Correlation),]
ggplot(data, aes(x=Robustness, y=Correlation, color=NumSplit, shape=pval_cat)) +
  geom_point() + theme_test() +
  labs(y="Correlation between number of stop codons\nand entropy of the distribution of reached peaks", x="Code robustness") +
  scale_color_viridis_d() + geom_hline(yintercept=0, linetype="dashed") +
  scale_shape_discrete(name="P-value")
```