---
title: "ParE2: Correlation of landscape properties with aaindex"
author: "Hana Rozhonova"
date: "6/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Read the data
```{r}
results = read.delim("output/results_aaindex", header=FALSE)
colnames(results) = c("Property", paste0("Corr_", c("numPeaks", "magEpi", "ssEpi", "rsEpi", "paths")))
bartonek_s01 = read.csv("input/Bartonek_PNAS_2020_S01.csv", header=TRUE)
colnames(bartonek_s01)[1] = "Property"
results = merge(results, bartonek_s01)
res_random_corrs = read.delim("output/results_aaindex_null", header=FALSE, sep=",")
colnames(res_random_corrs) = c("numPeaks", "magEpi", "ssEpi", "rsEpi", "paths")
```

# Peaks
### Number of peaks
P-value based on the randomly generated properties:
```{r}
results$P_numPeaks = sapply(results$Corr_numPeaks, function(x) (sum(abs(res_random_corrs[,1])>=abs(x))+1)/nrow(res_random_corrs))
```

Benjmaini-Hochberg correction for multiple testing:
```{r}
results$Padj_numPeaks = p.adjust(results$P_numPeaks, method="BH")
tmp = results[results$Padj_numPeaks<=0.05,]
nrow(tmp)
```

How many of the statistically significant properties are positively and negatively correlated?
```{r}
sum(tmp$Corr_numPeaks>=0)
sum(tmp$Corr_numPeaks<=0)
```


Which properties are statistically significant?
```{r}
summary(as.factor(tmp$category))
```

Is any category overrepresented?
```{r}
### "hydro"
# frequency in original data
(p=sum(results$category=="hydro")/nrow(results))
# frequency in the significant data
sum(tmp$category == "hydro")/nrow(tmp)
# binomial test
binom.test(sum(tmp$category == "hydro"), nrow(tmp), p, "g")

### alpha
# frequency in original data
(p = sum(results$category=="alpha")/nrow(results))
# frequency in the significant data
sum(tmp$category == "alpha")/nrow(tmp)
# binomial test
binom.test(sum(tmp$category == "alpha"), nrow(tmp), p, "g")

### beta
# frequency in original data
(p=sum(results$category=="beta")/nrow(results))
# frequency in the significant data
sum(tmp$category == "beta")/nrow(tmp)
# binomial test
binom.test(sum(tmp$category == "beta"), nrow(tmp), p, "g")

### other
# frequency in original data
(p=sum(results$category=="other")/nrow(results))
# frequency in the significant data
sum(tmp$category == "other")/nrow(tmp)
# binomial test
binom.test(sum(tmp$category == "other"), nrow(tmp), p, "g")
```



# Epistasis
### Magnitude epistasis
P-value based on the randomly generated properties:
```{r}
results$P_magEpi = sapply(results$Corr_magEpi, function(x) (sum(abs(res_random_corrs[,2])>=abs(x))+1)/nrow(res_random_corrs))
```

BH correction for multiple testing:
```{r}
results$Padj_magEpi = p.adjust(results$P_magEpi, method="BH")
tmp = results[results$Padj_magEpi<=0.05,]
nrow(tmp)
```

How many are positively and negatively correlated?
```{r}
sum(tmp$Corr_magEpi>=0)
sum(tmp$Corr_magEpi<=0)
```


Which properties are statistically significant?
```{r}
summary(as.factor(tmp$category))
```

Is any category overrepresented?
```{r}
### "hydro"
# frequency in original data
(p=sum(results$category=="hydro")/nrow(results))
# frequency in the significant data
sum(tmp$category == "hydro")/nrow(tmp)
# binomial test
binom.test(sum(tmp$category == "hydro"), nrow(tmp), p, "g")

### alpha
# frequency in original data
(p = sum(results$category=="alpha")/nrow(results))
# frequency in the significant data
sum(tmp$category == "alpha")/nrow(tmp)
# binomial test
binom.test(sum(tmp$category == "alpha"), nrow(tmp), p, "g")

### beta
# frequency in original data
(p=sum(results$category=="beta")/nrow(results))
# frequency in the significant data
sum(tmp$category == "beta")/nrow(tmp)
# binomial test
binom.test(sum(tmp$category == "beta"), nrow(tmp), p, "g")

### other
# frequency in original data
(p=sum(results$category=="other")/nrow(results))
# frequency in the significant data
sum(tmp$category == "other")/nrow(tmp)
# binomial test
binom.test(sum(tmp$category == "other"), nrow(tmp), p, "g")
```


### Simple-sign epistasis
P-value based on the randomly generated properties:
```{r}
results$P_ssEpi = sapply(results$Corr_ssEpi, function(x) (sum(abs(res_random_corrs[,3])>=abs(x))+1)/nrow(res_random_corrs))
```

BH correction for multiple testing:
```{r}
results$Padj_ssEpi = p.adjust(results$P_ssEpi, method="BH")
tmp = results[results$Padj_ssEpi<=0.05,]
nrow(tmp)
```

How many are positively and negatively correlated?
```{r}
sum(tmp$Corr_ssEpi>=0)
sum(tmp$Corr_ssEpi<=0)
```


Which properties are statistically significant?
```{r}
summary(as.factor(tmp$category))
```

Is any category overrepresented?
```{r}
### "hydro"
# frequency in original data
(p=sum(results$category=="hydro")/nrow(results))
# frequency in the significant data
sum(tmp$category == "hydro")/nrow(tmp)
# binomial test
binom.test(sum(tmp$category == "hydro"), nrow(tmp), p, "g")

### alpha
# frequency in original data
(p = sum(results$category=="alpha")/nrow(results))
# frequency in the significant data
sum(tmp$category == "alpha")/nrow(tmp)
# binomial test
binom.test(sum(tmp$category == "alpha"), nrow(tmp), p, "g")

### beta
# frequency in original data
(p=sum(results$category=="beta")/nrow(results))
# frequency in the significant data
sum(tmp$category == "beta")/nrow(tmp)
# binomial test
binom.test(sum(tmp$category == "beta"), nrow(tmp), p, "g")

### other
# frequency in original data
(p=sum(results$category=="other")/nrow(results))
# frequency in the significant data
sum(tmp$category == "other")/nrow(tmp)
# binomial test
binom.test(sum(tmp$category == "other"), nrow(tmp), p, "g")
```


### Reciprocal-sign epistasis
P-value based on the randomly generated properties:
```{r}
results$P_rsEpi = sapply(results$Corr_rsEpi, function(x) (sum(abs(res_random_corrs[,4])>=abs(x))+1)/nrow(res_random_corrs))
```

BH correction for multiple testing:
```{r}
results$Padj_rsEpi = p.adjust(results$P_rsEpi, method="BH")
tmp = results[results$Padj_rsEpi<=0.05,]
nrow(tmp)
```

How many are positively and negatively correlated?
```{r}
sum(tmp$Corr_rsEpi>=0)
sum(tmp$Corr_rsEpi<=0)
```


Which properties are statistically significant?
```{r}
summary(as.factor(tmp$category))
```

Is any category overrepresented?
```{r}
### "hydro"
# frequency in original data
(p=sum(results$category=="hydro")/nrow(results))
# frequency in the significant data
sum(tmp$category == "hydro")/nrow(tmp)
# binomial test
binom.test(sum(tmp$category == "hydro"), nrow(tmp), p, "g")

### alpha
# frequency in original data
(p = sum(results$category=="alpha")/nrow(results))
# frequency in the significant data
sum(tmp$category == "alpha")/nrow(tmp)
# binomial test
binom.test(sum(tmp$category == "alpha"), nrow(tmp), p, "g")

### beta
# frequency in original data
(p=sum(results$category=="beta")/nrow(results))
# frequency in the significant data
sum(tmp$category == "beta")/nrow(tmp)
# binomial test
binom.test(sum(tmp$category == "beta"), nrow(tmp), p, "g")

### other
# frequency in original data
(p=sum(results$category=="other")/nrow(results))
# frequency in the significant data
sum(tmp$category == "other")/nrow(tmp)
# binomial test
binom.test(sum(tmp$category == "other"), nrow(tmp), p, "g")
```


# Accessibility of the global peak
P-value based on the randomly generated properties:
```{r}
results$P_paths = sapply(results$Corr_paths, function(x) (sum(abs(res_random_corrs[,5])>=abs(x))+1)/nrow(res_random_corrs))
```

BH correction for multiple testing:
```{r}
results$Padj_paths = p.adjust(results$P_paths, method="BH")
tmp = results[results$Padj_paths<=0.05,]
nrow(tmp)
```

How many are positively and negatively correlated?
```{r}
sum(tmp$Corr_paths>=0)
sum(tmp$Corr_paths<=0)
```


Which properties are statistically significant?
```{r}
summary(as.factor(tmp$category))
```

Is any category overrepresented?
```{r}
### "hydro"
# frequency in original data
(p=sum(results$category=="hydro")/nrow(results))
# frequency in the significant data
sum(tmp$category == "hydro")/nrow(tmp)
# binomial test
binom.test(sum(tmp$category == "hydro"), nrow(tmp), p, "g")

### alpha
# frequency in original data
(p = sum(results$category=="alpha")/nrow(results))
# frequency in the significant data
sum(tmp$category == "alpha")/nrow(tmp)
# binomial test
binom.test(sum(tmp$category == "alpha"), nrow(tmp), p, "g")

### beta
# frequency in original data
(p=sum(results$category=="beta")/nrow(results))
# frequency in the significant data
sum(tmp$category == "beta")/nrow(tmp)
# binomial test
binom.test(sum(tmp$category == "beta"), nrow(tmp), p, "g")

### other
# frequency in original data
(p=sum(results$category=="other")/nrow(results))
# frequency in the significant data
sum(tmp$category == "other")/nrow(tmp)
# binomial test
binom.test(sum(tmp$category == "other"), nrow(tmp), p, "g")
```



# Save the data
```{r}
write.table(results[,c(1,2,22,23,3,24,25,4,26,27,5,28,29,6,30,31,7)], file="output/supp_data_S2.csv", quote=FALSE, sep=",", row.names=FALSE)
```

