---
title: "Modelling GB1 and the ParD3 landscapes using an additive model"
author: "Hana"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12)
```

```{r}
aas = c("A", "C", "D", "E", "F", "G", "H", "I", "K", "L", "M", "N", "P", "Q", "R", "S", "T", "V", "W", "Y")
```



# GB1
Read the data:
```{r}
data_gb1 = read.csv("../../GB1/01_vcregression/output/map.txt", stringsAsFactors = FALSE)
```

Create a data frame with 80 dummy variables (4 positions x 20 amino acis) and the scores:
```{r}
df_gb1 = data.frame(matrix(0, nrow=nrow(data_gb1), ncol=81))
colnames(df_gb1) = c("Score", paste0(rep(c(1,2,3,4), each=20), rep(c("A", "C", "D", "E", "F", "G", "H", "I", "K", "L", "M", "N", "P", "Q", "R", "S", "T", "V", "W", "Y"), 4)))

for (i in 1:nrow(data_gb1)) {
  genotype = data_gb1$sequence[i]
  df_gb1$Score[i] = data_gb1$phenotype[i]
  for (j in 1:4) {
    col = paste0(j, substr(genotype, j, j))
    df_gb1[i,col] = 1
  }
}
```

Fit a linear model:
```{r}
model_gb1 = lm(Score ~ ., data=df_gb1)
summary(model_gb1)
```

The linear model explains ca 52.5 % of variance.



# ParD3
## ParE3
Read the data:
```{r}
data_pare3 = read.csv("../../ParE3/01_vcregression/output/map.txt", stringsAsFactors = FALSE)
```

Create a data frame with 80 dummy variables (3 positions x 20 amino acis) and the scores:
```{r}
df_pare3 = data.frame(matrix(0, nrow=nrow(data_pare3), ncol=61))
colnames(df_pare3) = c("Score", paste0(rep(c(1,2,3), each=20), rep(c("A", "C", "D", "E", "F", "G", "H", "I", "K", "L", "M", "N", "P", "Q", "R", "S", "T", "V", "W", "Y"), 3)))

for (i in 1:nrow(data_pare3)) {
  genotype = data_pare3$sequence[i]
  df_pare3$Score[i] = data_pare3$phenotype[i]
  for (j in 1:3) {
    col = paste0(j, substr(genotype, j, j))
    df_pare3[i,col] = 1
  }
}
```

Fit a linear model:
```{r}
model_pare3 = lm(Score ~ ., data=df_pare3)
summary(model_pare3)
```

The linear model explains ca 85.6 % of variance.


## ParE2
Read the data:
```{r}
data_pare2 = read.csv("../../ParE2/01_vcregression/output/map.txt", stringsAsFactors = FALSE)
```

Create a data frame with 80 dummy variables (3 positions x 20 amino acis) and the scores:
```{r}
df_pare2 = data.frame(matrix(0, nrow=nrow(data_pare2), ncol=61))
colnames(df_pare2) = c("Score", paste0(rep(c(1,2,3), each=20), rep(c("A", "C", "D", "E", "F", "G", "H", "I", "K", "L", "M", "N", "P", "Q", "R", "S", "T", "V", "W", "Y"), 3)))

for (i in 1:nrow(data_pare2)) {
  genotype = data_pare2$sequence[i]
  df_pare2$Score[i] = data_pare2$phenotype[i]
  for (j in 1:3) {
    col = paste0(j, substr(genotype, j, j))
    df_pare2[i,col] = 1
  }
}
```

Fit a linear model:
```{r}
model_pare2 = lm(Score ~ ., data=df_pare2)
summary(model_pare2)
```

The linear model explains ca 84.5 % of variance.

# ParB
## ParS
Read the data:
```{r}
data_parS = read.csv("../../ParS/01_vcregression/output/map.txt", stringsAsFactors = FALSE)
```

Create a data frame with 80 dummy variables (4 positions x 20 amino acis) and the scores:
```{r}
df_parS = data.frame(matrix(0, nrow=nrow(data_parS), ncol=81))
colnames(df_parS) = c("Score", paste0(rep(c(1,2,3,4), each=20), rep(c("A", "C", "D", "E", "F", "G", "H", "I", "K", "L", "M", "N", "P", "Q", "R", "S", "T", "V", "W", "Y"), 4)))

for (i in 1:nrow(data_parS)) {
  genotype = data_parS$sequence[i]
  df_parS$Score[i] = data_parS$phenotype[i]
  for (j in 1:4) {
    col = paste0(j, substr(genotype, j, j))
    df_parS[i,col] = 1
  }
}
```

Fit a linear model:
```{r}
model_parS = lm(Score ~ ., data=df_parS)
summary(model_parS)
```

The linear model explains ca 66.2 % of variance.

## NBS
Read the data:
```{r}
data_NBS = read.csv("../../NBS/01_vcregression/output/map.txt", stringsAsFactors = FALSE)
```

Create a data frame with 80 dummy variables (4 positions x 20 amino acis) and the scores:
```{r}
df_NBS = data.frame(matrix(0, nrow=nrow(data_NBS), ncol=81))
colnames(df_NBS) = c("Score", paste0(rep(c(1,2,3,4), each=20), rep(c("A", "C", "D", "E", "F", "G", "H", "I", "K", "L", "M", "N", "P", "Q", "R", "S", "T", "V", "W", "Y"), 4)))

for (i in 1:nrow(data_NBS)) {
  genotype = data_NBS$sequence[i]
  df_NBS$Score[i] = data_NBS$phenotype[i]
  for (j in 1:4) {
    col = paste0(j, substr(genotype, j, j))
    df_NBS[i,col] = 1
  }
}
```

Fit a linear model:
```{r}
model_NBS = lm(Score ~ ., data=df_NBS)
summary(model_NBS)
```

The linear model explains ca 52.5 % of variance.