---
title: "ParE3, greedy walks, aa permutation codes"
author: "Hana Rozhonova"
date: "6/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(stringr)
library(ggplot2)
library(gridExtra)
library(ggpubr)
library(dplyr)
library(reticulate)
```

# Read the data
```{r}
if (file.exists("output/results_processed.tsv")) {
  results = read.delim("output/results_processed.tsv", header=TRUE)
} else {
  results = read.delim("output/results", header=FALSE, stringsAsFactors = FALSE)
  colnames(results) = c("Seed","SplitAa", "Robustness", "MeanFitness", "MeanSteps", "Peaks")
  
  # entropy of the reached peaks
  entropy_peaks = c()
  for (i in 1:nrow(results)) {
    peaks_counts = str_split(results$Peaks[i], ",")[[1]]
    peaks_counts = head(peaks_counts, -1)
    peaks_counts = as.numeric(lapply(str_split(peaks_counts, ":"), function(x) x[2]))
    # compute Shannon entropy
    peaks_probs = peaks_counts/sum(peaks_counts)
    entropy = -1*sum(peaks_probs*log(peaks_probs))
    entropy_peaks = c(entropy_peaks, entropy)
  }
  results$Entropy_peaks = entropy_peaks
  
  # probability of reaching the global peak
  prop = function(peaks, opt) {
    peaks_split = strsplit(peaks, ",")[[1]]
    peak_opt = peaks_split[startsWith(peaks_split, opt)]
    peak_opt_count = sum(as.numeric(unlist(strsplit(peak_opt, ":"))[seq(2, 2*length(peak_opt), 2)]))
    count_total = 0
    for (peak in peaks_split) {
      count_total = count_total + as.numeric(strsplit(peak, ":")[[1]][2])
    }
    return(peak_opt_count/count_total)
  }
  # global max
  data = read.delim("input/map.tsv", header=TRUE)
  global_max = as.character(data$sequence[which.max(data$phenotype)])
  results$Prob_reaching_global_peak = sapply(results$Peaks, prop, global_max)
  write.table(results, file="output/results_processed.tsv", quote=FALSE, sep="\t", row.names=FALSE)
}
```

# Histograms: Mean fitness, mean number of steps, entropy of the reached peaks, and probability of reaching the global peak(s)
```{r}
fit_standard = results$MeanFitness[results$Seed==0]
p = sum(results$MeanFitness>=fit_standard)/nrow(results)
hist_fit = ggplot(results, aes(x=MeanFitness)) + 
  geom_histogram(alpha=0.2, color="black") +
  theme_test() +
  geom_vline(xintercept=fit_standard, linetype="dashed", color="red") +
  labs(x="Mean fitness", y="Count", title=paste0("p=", signif(p, digits=3)))

steps_standard = results$MeanSteps[results$Seed==0]
p = sum(results$MeanSteps>=steps_standard)/nrow(results)
hist_steps = ggplot(results, aes(x=MeanSteps)) + 
  geom_histogram(alpha=0.2, color="black") +
  theme_test() +
  geom_vline(xintercept=steps_standard, linetype="dashed", color="red") +
  labs(x="Mean number of steps", y="Count", title=paste0("p=", signif(p, digits=3)))

ent_standard = results$Entropy_peaks[results$Seed==0]
p = sum(results$Entropy_peaks<=ent_standard)/nrow(results)
hist_ent = ggplot(results, aes(x=Entropy_peaks)) + 
  geom_histogram(alpha=0.2, color="black") +
  theme_test() +
  geom_vline(xintercept=ent_standard, linetype="dashed", color="red") +
  labs(x="Entropy of the distribution of reached peaks", y="Count", title=paste0("p=", signif(p, digits=3)))

prob_standard = results$Prob_reaching_global_peak[results$Seed==0]
p = sum(results$Prob_reaching_global_peak>=prob_standard)/nrow(results)
hist_prob = ggplot(results, aes(x=Prob_reaching_global_peak)) + 
  geom_histogram(alpha=0.2, color="black") +
  theme_test() +
  geom_vline(xintercept=prob_standard, linetype="dashed", color="red") +
  labs(x="Probability of reaching the global peak(s)", y="Count", title=paste0("p=", signif(p, digits=3)))

grid.arrange(hist_fit, hist_steps, hist_ent, hist_prob, ncol=2)
```


# Correlations with code robustness
```{r}
rob_standard = results$Robustness[results$Seed==0]

(corTest = cor.test(results$MeanFitness, results$Robustness))
p_fit = ggplot(results, aes(x=Robustness, y=MeanFitness)) +
  geom_bin2d(bins=59, alpha=0.8) +
  scale_fill_continuous(type = "viridis", name="Count") +
  theme_test() +
  theme(legend.position="bottom") +
  geom_point(x=rob_standard, y=fit_standard, fill="gray", pch=21, size=2) + 
  labs(x="Code robustness", y="Mean fitness reached\nby greedy adaptive walks") +
  geom_smooth(color="black") +
  geom_text(label=paste0("R = ", signif(corTest$estimate, digits=3)), x=0.43, y=min(results$MeanFitness)+0.03*(max(results$MeanFitness)-min(results$MeanFitness)), size=4) +
  annotate(geom = "label", x=rob_standard, y=fit_standard+0.07*(max(results$MeanFitness)-min(results$MeanFitness)), label="SGC", alpha=0.8, size=3)

(corTest = cor.test(results$MeanSteps, results$Robustness))
p_steps =  ggplot(results, aes(x=Robustness, y=MeanSteps)) +
  geom_bin2d(bins=59, alpha=0.8) +
  scale_fill_continuous(type = "viridis", name="Count") +
  theme_test() +
  theme(legend.position="bottom") +
  geom_point(x=rob_standard, y=steps_standard, fill="gray", pch=21, size=2) + 
  labs(x="Code robustness", y="Mean number of steps\nof greedy adaptive walks") +
  geom_smooth(color="black") +
  geom_text(label=paste0("R = ", signif(corTest$estimate, digits=3)), x=0.43, y=min(results$MeanSteps)+0.03*(max(results$MeanSteps)-min(results$MeanSteps)), size=4) +
  annotate(geom = "label", x=rob_standard, y=steps_standard+0.07*(max(results$MeanSteps)-min(results$MeanSteps)), label="SGC", alpha=0.8, size=3)


(corTest = cor.test(results$Entropy_peaks, results$Robustness))
p_ent = ggplot(results, aes(x=Robustness, y=Entropy_peaks)) +
  geom_bin2d(bins=59, alpha=0.8) +
  scale_fill_continuous(type = "viridis", name="Count") +
  theme_test() +
  theme(legend.position="bottom") +
  geom_point(x=rob_standard, y=ent_standard, fill="gray", pch=21, size=2) + 
  labs(x="Code robustness", y="Entropy of the distribution\nof reached peaks") +
  geom_smooth(color="black") +
  geom_text(label=paste0("R = ", signif(corTest$estimate, digits=3)), x=0.43, y=min(results$Entropy_peaks)+0.03*(max(results$Entropy_peaks)-min(results$Entropy_peaks)), size=4) +
  annotate(geom = "label", x=rob_standard, y=ent_standard+0.07*(max(results$Entropy_peaks)-min(results$Entropy_peaks)), label="SGC", alpha=0.8, size=3)

(corTest = cor.test(results$Prob_reaching_global_peak, results$Robustness))
p_prob = ggplot(results, aes(x=Robustness, y=Prob_reaching_global_peak)) +
  geom_bin2d(bins=59) +
  scale_fill_continuous(type = "viridis") +
  theme_test() +
  geom_point(x=rob_standard, y=prob_standard, color="red") + 
  labs(x="Code robustness", y="Probability of reaching the global peak", title=paste0("R=", signif(corTest$estimate, digits=3)))

grid.arrange(p_fit, p_steps, p_ent, p_prob, nrow=2)
```

Restrict only to codes that have the same size of the global peak as the SGC and the global peak is not split:
```{r}
res_summitSizes = read.delim("../..//02_ruggedness/aa_permutation/02_01_global_peak_size/results_globalPeakSize_aaPerm", header=FALSE)
colnames(res_summitSizes) = c("Seed", "SummitSize")
results = merge(results, res_summitSizes)

cor.test(results$MeanFitness, results$SummitSize)

tmp = results[results$SummitSize == results$SummitSize[results$Seed==0] & ! results$SplitAa %in% c("E", "D", "W"),]


(corTest = cor.test(tmp$MeanFitness, tmp$Robustness))
(corTest = cor.test(tmp$MeanSteps, tmp$Robustness))
(corTest = cor.test(tmp$Entropy_peaks, tmp$Robustness))
(corTest = cor.test(tmp$Prob_reaching_global_peak, tmp$Robustness))
```


What drives the correlation between code robustness and mean fitness? 

(1) Discarding the global peak -- is the correlation still present?
```{r}
data_gb1=read.delim("../../01_vcregression/output/map.txt", header=TRUE, sep=",")
max_fitness = max(data_gb1$phenotype)
tmp$MeanFitness_woSummit = 1/(1-tmp$Prob_reaching_global_peak)*(tmp$MeanFitness - tmp$Prob_reaching_global_peak*max_fitness)
cor.test(tmp$Robustness, tmp$MeanFitness_woSummit)
```

The correlation is not driven by the summit.

(2) Is it because the local peaks in the robust landscapes are higher than the local peaks in less robust landscapes? 
```{r}
res_ruggedness= read.delim("../../02_ruggedness/aa_permutation/output/results/results", header=FALSE)
colnames(res_ruggedness) = c("Seed", "SplitAa", "Robustness", "NumPeaks", "MeanPeakHeight", "NoEpi", "MagEpi", "SsEpi", "RsEpi", "Accessibility")
tmp = merge(tmp, res_ruggedness[,c("Seed", "MeanPeakHeight")])
cor.test(tmp$Robustness, tmp$MeanPeakHeight)
```

(3) So it must be that under the robust codes the basin of attraction of higher peaks is relatively larger than for less robust codes.
```{r}
write.table(tmp, "output/results_constSummitSize.tsv", quote=FALSE, sep="\t", row.names=FALSE)
py_run_file("data_basins.py")
res_basins = read.delim("output/basins.tsv", header=FALSE)
colnames(res_basins) = c("Seed", "Robustness", "sequence", "Basin")
res_basins = merge(res_basins, data_gb1)

coefs = c()
robs = c()
for (seed in unique(res_basins$Seed)) {
  tmp_basins = res_basins[res_basins$Seed==seed,]
  lm = lm(log(Basin)~phenotype, data=tmp_basins)
  coef = lm$coefficients[2]
  coefs = c(coefs, coef)
  robs = c(robs, tmp$Robustness[tmp$Seed==seed])
}
cor.test(robs, coefs)
```



# The final plot
```{r}
ggarrange(p_fit, p_ent, p_steps,
          labels = c("A", "B", "C"),
          ncol=3)
```

