---
title: "GB1, landscape ruggedness, aa permutation codes"
author: "Hana Rozhonova"
date: "5/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(gridExtra)
library(stringr)
library(ggpubr)
```

# Read the data
```{r}
results = read.delim("output/results/results", header=FALSE)
colnames(results) = c("Seed", "SplitAa", "Robustness", "NumPeaks", "MeanPeakHeight", "NoEpi", "MagEpi", "SsEpi", "RsEpi", "PropAccPaths")
```

# Peaks
#### Histogram of the number of peaks
```{r}
peaks_standard = results$NumPeaks[results$Seed==0]
p = sum(results$NumPeaks<=peaks_standard)/nrow(results)
ggplot(results, aes(x=NumPeaks)) + 
  geom_histogram(alpha=0.2, color="black") +
  theme_test() +
  geom_vline(xintercept=peaks_standard, linetype="dashed", color="red") +
  labs(x="Number of peaks", y="Count", title=paste0("p=", signif(p, digits=3)))
```

#### Correlation of number of peaks with code robustness
```{r}
(corTest = cor.test(results$NumPeaks, results$Robustness))
```


Density plot:
```{r}
rob_standard = results$Robustness[results$Seed==0]
(p_peaks = ggplot(results, aes(x=Robustness, y=NumPeaks)) +
  geom_bin2d(bins=59, alpha=0.8) +
  scale_fill_continuous(type = "viridis", name="Count") +
  theme_test() +
  theme(legend.position="bottom") +
  geom_point(x=rob_standard, y=peaks_standard, fill="gray", pch=21, size=2) + 
  labs(x="Code robustness", y="Number of adaptive peaks") +
  geom_smooth(color="black") +
  geom_text(label=paste0("R = ", signif(corTest$estimate, digits=3)), x=0.43, y=max(results$NumPeaks)-0.03*(max(results$NumPeaks)-min(results$NumPeaks)), size=4) +
  annotate(geom = "label", x=rob_standard, y=peaks_standard+0.07*(max(results$NumPeaks)-min(results$NumPeaks)), label="SGC", alpha=0.8, size=3))
```


#### Number of peaks vs. mean height of peak
```{r}
meanHeight_standard = results$MeanPeakHeight[results$Seed==0]

corTest = cor.test(results$NumPeaks, results$MeanPeakHeight)

ggplot(results, aes(x=NumPeaks, y=MeanPeakHeight)) +
  geom_bin2d(bins=59) +
  scale_fill_continuous(type = "viridis")  +
  theme_test() +
  geom_point(y=meanHeight_standard, x=peaks_standard, color="red") +
  labs(y="Mean peak height", x="Number of peaks", title=paste0("R=", signif(corTest$estimate, digits=3)))
```

Correlation of code robustness with mean peak height:
```{r}
(corTest = cor.test(results$Robustness, results$MeanPeakHeight))

ggplot(results, aes(x=Robustness, y=MeanPeakHeight)) +
  geom_bin2d(bins=59) +
  scale_fill_continuous(type = "viridis")  +
  theme_test() +
  geom_point(y=meanHeight_standard, x=rob_standard, color="red") +
  labs(y="Mean peak height", x="Code robustness", title=paste0("R=", signif(corTest$estimate, digits=3)))
```



# Accessibility of the global peak
#### Histogram
```{r}
propAccPaths_standard = results$PropAccPaths[results$Seed==0]
p = sum(results$PropAccPaths>=propAccPaths_standard)/nrow(results)
ggplot(results, aes(x=PropAccPaths)) + 
  geom_histogram(alpha=0.2, color="black") +
  theme_test() +
  geom_vline(xintercept=propAccPaths_standard, linetype="dashed", color="red") +
  labs(x="Probability that a randomly chosen path to the summit is accessible", y="Count", title=paste0("p=", signif(p, digits=3)))

```

#### Correlation with code robustness
```{r}
(corTest = cor.test(results$Robustness, results$PropAccPaths))
```

Correlation between peak accessibility and the size of the global peak:
```{r}
res_size = read.delim("02_01_global_peak_size/results_globalPeakSize_aaPerm", header=FALSE)
colnames(res_size) = c("Seed", "SummitSize")
results = merge(results, res_size)

cor.test(results$SummitSize, results$PropAccPaths)
size_standard = results$SummitSize[results$Seed==0]

ggplot(results, aes(x=SummitSize, y=PropAccPaths, group=SummitSize)) +
  geom_boxplot(notch=FALSE, fill="gray80", outlier.size=0.1, lwd=0.2)+
  scale_x_log10() +
  theme_test() +
  labs(x="Number of mRNAs in the summit", y="Probability that a randomly chosen path\nto the summit is accessible", title="GB1") 
```

Is summit accessibility significantly bigger if the peak is split?
```{r}
results$SplitSummit = results$SplitAa %in% c("W", "L", "A")
t.test(PropAccPaths ~ SplitSummit, data=results)
ggplot(results, aes(x=SplitSummit, y=PropAccPaths)) + 
  geom_boxplot(fill="gray80", outlier.size=0.5) +
  theme_test() +
  labs(x="Split summit", y="Probability that a randomly chosen path\nto the summit is accessible", title="GB1") 
```

Only restrict the analysis to those where the size of the global peak is the same as in the SGC and the summit is not split:
```{r}
tmp_res = results[results$SummitSize==results$SummitSize[results$Seed==0] & !results$SplitAa %in% c("W", "L", "A"),]
nrow(tmp_res)

(corTest = cor.test(tmp_res$Robustness, tmp_res$PropAccPaths))
```

The plot:
```{r}
(p_paths = ggplot(tmp_res, aes(x=Robustness, y=PropAccPaths)) +
  geom_bin2d(bins=50, alpha=0.8) +
  scale_fill_continuous(type = "viridis", name="Count") +
  theme_test() +
   theme(legend.position="bottom") +
  geom_point(x=rob_standard, y=propAccPaths_standard, fill="gray", pch=21, size=2) + 
  labs(x="Code robustness", y="Accessibility of the global peak") +
  geom_smooth(color="black") +
  geom_text(label=paste0("R = ", signif(corTest$estimate, digits=3)), x=0.42, y=max(tmp_res$PropAccPaths)-0.03*(max(tmp_res$PropAccPaths)-min(tmp_res$PropAccPaths)), size=4) +
  annotate(geom = "label", x=rob_standard, y=propAccPaths_standard-0.07*(max(tmp_res$PropAccPaths)-min(tmp_res$PropAccPaths)), label="SGC", alpha=0.8, size=3))

```



# Epistasis
#### Histograms
```{r}
# magnitude
magEpi_standard = results$MagEpi[results$Seed==0]
p = sum(results$MagEpi<=magEpi_standard)/nrow(results)
p_m = ggplot(results, aes(x=MagEpi)) + 
  geom_histogram(alpha=0.2, color="black") +
  theme_test() +
  geom_vline(xintercept=magEpi_standard, linetype="dashed", color="red") +
  labs(x="Proportion of squares with magnitude epistasis", y="Count", title=paste0("p=", signif(p, digits=3)))

# simple sign
ssEpi_standard = results$SsEpi[results$Seed==0]
p = sum(results$SsEpi<=ssEpi_standard)/nrow(results)
p_ss = ggplot(results, aes(x=SsEpi)) + 
  geom_histogram(alpha=0.2, color="black") +
  theme_test() +
  geom_vline(xintercept=ssEpi_standard, linetype="dashed", color="red") +
  labs(x="Proportion of squares with simple-sign epistasis", y="Count", title=paste0("p=", signif(p, digits=3)))

# reciprocal sign
rsEpi_standard = results$RsEpi[results$Seed==0]
p = sum(results$RsEpi<=rsEpi_standard)/nrow(results)
p_rs = ggplot(results, aes(x=RsEpi)) + 
  geom_histogram(alpha=0.2, color="black") +
  theme_test() +
  geom_vline(xintercept=rsEpi_standard, linetype="dashed", color="red") +
  labs(x="Proportion of squares with reciprocal-sign epistasis", y="Count", title=paste0("p=", signif(p, digits=3)))

grid.arrange(p_m, p_ss, p_rs, ncol=3)
```


#### Correlation with code robustness
```{r}
cor.test(results$Robustness, results$MagEpi)

cor.test(results$Robustness, results$SsEpi)

(corTest = cor.test(results$Robustness, results$RsEpi))
p_rsEpi = ggplot(results, aes(x=Robustness, y=RsEpi)) +
  geom_bin2d(bins=59, alpha=0.8) +
  scale_fill_continuous(type = "viridis", name="Count") +
  theme_test() +
  theme(legend.position="bottom") +
  geom_point(x=rob_standard, y=rsEpi_standard, fill="gray", pch=21, size=2) + 
  labs(x="Code robustness", y="Prevalence of reciprocal-sign\nepistasis") +
  geom_smooth(color="black") +
  #theme(legend.position="none") +
  geom_text(label=paste0("R = ", signif(corTest$estimate, digits=3)), x=0.43, y=max(results$RsEpi)-0.03*(max(results$RsEpi)-min(results$RsEpi)), size=4) +
  annotate(geom = "label", x=rob_standard, y=rsEpi_standard-0.07*(max(results$RsEpi)-min(results$RsEpi)), label="SGC", alpha=0.8, size=3)
```



# The final plot
```{r}
ggarrange(p_peaks,p_rsEpi, p_paths,
          labels = c("A", "B", "C"),
          ncol=3)
```

