---
title: "Most and least evolvable Ostrov codes"
output: html_document
date: "2022-10-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggvenn)
library(ggplot2)
library(viridisLite)
library(ggpubr)
```

# Read the greedy simulation results
```{r}
res_gb1 = read.delim("../../GB1/03_greedy_walks/ostrov/output/results_processed.tsv")
res_pare2 = read.delim("../../ParE2/03_greedy_walks/ostrov/output/results_processed.tsv")
res_pare3 = read.delim("../../ParE3/03_greedy_walks/ostrov/output/results_processed.tsv")
```


# Correlation between code performance on different data sets
```{r}
cor.test(res_gb1$MeanFitness, res_pare2$MeanFitness)
cor.test(res_gb1$MeanFitness, res_pare3$MeanFitness)
cor.test(res_pare3$MeanFitness, res_pare2$MeanFitness)
```

# The best 20% of codes, in terms of fitness, for each data set
```{r}
res_gb1 = res_gb1[order(res_gb1$MeanFitness),]
bestCodes_gb1 = res_gb1[ceiling(0.8*194481):194481,]
res_pare2 = res_pare2[order(res_pare2$MeanFitness),]
bestCodes_pare2 = res_pare2[ceiling(0.8*194481):194481,]
res_pare3 = res_pare3[order(res_pare3$MeanFitness),]
bestCodes_pare3 = res_pare3[ceiling(0.8*194481):194481,]
```

# Venn diagram of the best 20% of codes
```{r}
x <- list(
  GB1 = bestCodes_gb1$Id,
  ParE2 = bestCodes_pare2$Id,
  ParE3 = bestCodes_pare3$Id
  )
p_venn = ggvenn(
  x, 
  fill_color = viridis(3),
  stroke_size = 0.5, set_name_size = 4, text_size=2.7
  )
```

# Properties of the most evolvable codes
```{r}
bestCodes = intersect(intersect(bestCodes_gb1$Id, bestCodes_pare2$Id), bestCodes_pare3$Id)
res_gb1$BestCode = res_gb1$Id %in% bestCodes
res_gb1$BestCode = ifelse(res_gb1$BestCode==TRUE, "Consistently high-ranking", "Remaining")
```

## Robustness
```{r}
t.test(Robustness ~ BestCode, res_gb1)


p_rob = ggplot(res_gb1, aes(x=Robustness, fill=BestCode, color=BestCode)) +
  facet_grid(rows=vars(BestCode)) + 
  geom_histogram(aes(y=..density..), alpha=0.5, bins=23) +
  theme_test() +
  scale_color_manual(values=c("grey", "#afb860"), name="") +
  scale_fill_manual(values=c("grey", "#afb860"), name="") +
  labs(x="Code robustness", y="Density") +
  theme(
  strip.background = element_blank(),
  strip.text.y = element_blank()
)
```

## Number of split codon blocks
```{r}
t.test(numSplitBlocks ~ BestCode, res_gb1)

p_numSplit = ggplot(res_gb1, aes(x=numSplitBlocks, fill=BestCode, color=BestCode)) +
  facet_grid(rows=vars(BestCode)) + 
  geom_histogram(aes(y=..density..), alpha=0.5, bins=5) +
  theme_test() +
  scale_color_manual(values=c("grey", "#afb860"), name="") +
  scale_fill_manual(values=c("grey", "#afb860"), name="") +
  labs(x="Number of split codon blocks", y="Density") +
  theme(strip.background = element_blank(), strip.text.y = element_blank()
)
```

## Number of stop codons
```{r}
t.test(numStop ~ BestCode, res_gb1)


p_numStop = ggplot(res_gb1, aes(x=numStop, fill=BestCode, color=BestCode)) +
  facet_grid(rows=vars(BestCode)) + 
  geom_histogram(aes(y=..density..), alpha=0.5, bins=8) +
  theme_test() +
  scale_color_manual(values=c("grey", "#afb860"), name="") +
  scale_fill_manual(values=c("grey", "#afb860"), name="") +
  labs(x="Number of stop codons", y="Density") +
  theme(strip.background = element_blank(), strip.text.y = element_blank()) +
  scale_x_continuous(breaks=2:9)
```

## The final plot
```{r}
ggarrange(p_venn, 
          ggarrange(p_rob, p_numSplit, p_numStop,
          ncol=3, labels=c("B", "C", "D"), common.legend=TRUE, legend="bottom"),
          ncol=2, labels=c("A", ""), widths=c(1.5,3))
```

# Well-performing codes
```{r}
# the best code - in terms of robustness, number of split codon blocks, number of stop codons
res_mostRobust = res_gb1[res_gb1$Robustness==max(res_gb1$Robustness) &
                           res_gb1$BestCode=="Consistently high-ranking",]
res_gb1[res_gb1$Robustness==max(res_gb1$Robustness) &
          res_gb1$numSplitBlocks==0 &
          res_gb1$numStop==2 &
          res_gb1$BestCode=="Consistently high-ranking",1:10]

# the best codes that would only reassign UAG
tmp = res_gb1[ res_gb1$Block1=="-" & res_gb1$Block3=="X" & res_gb1$Block4=="-" & res_gb1$Block2!="-" ,]
tmp[tmp$BestCode=="Consistently high-ranking",1:10]
```


# Least evolvable codes
## The worst 20% of codes, in terms of fitness, for each data set
```{r}
res_gb1 = res_gb1[order(res_gb1$MeanFitness),]
worstCodes_gb1 = res_gb1[1:ceiling(0.2*194481),]
res_pare2 = res_pare2[order(res_pare2$MeanFitness),]
worstCodes_pare2 = res_pare2[1:ceiling(0.2*194481),]
res_pare3 = res_pare3[order(res_pare3$MeanFitness),]
worstCodes_pare3 = res_pare3[1:ceiling(0.2*194481),]
```

## Venn diagram of the worst 20% of codes
```{r}
x <- list(
  GB1 = worstCodes_gb1$Id,
  ParE2 = worstCodes_pare2$Id,
  ParE3 = worstCodes_pare3$Id
  )
p_venn = ggvenn(
  x, 
  fill_color = viridis(3),
  stroke_size = 0.5, set_name_size = 4, text_size=2.7
  )
```

## Properties of the least evolvable codes
```{r}
worstCodes = intersect(intersect(worstCodes_gb1$Id, worstCodes_pare2$Id), worstCodes_pare3$Id)
res_gb1$WorstCode = res_gb1$Id %in% worstCodes
res_gb1$WorstCode = ifelse(res_gb1$WorstCode==TRUE, "Consistently low-ranking", "Remaining")
```

## Robustness
```{r}
t.test(Robustness ~ WorstCode, res_gb1)


p_rob = ggplot(res_gb1, aes(x=Robustness, fill=WorstCode, color=WorstCode)) +
  facet_grid(rows=vars(WorstCode)) + 
  geom_histogram(aes(y=..density..), alpha=0.5, bins=23) +
  theme_test() +
  scale_color_manual(values=c("grey", "#afb860"), name="") +
  scale_fill_manual(values=c("grey", "#afb860"), name="") +
  labs(x="Code robustness", y="Density") +
  theme(
  strip.background = element_blank(),
  strip.text.y = element_blank()
)
```

## Number of split codon blocks
```{r}
t.test(numSplitBlocks ~ WorstCode, res_gb1)

p_numSplit = ggplot(res_gb1, aes(x=numSplitBlocks, fill=WorstCode, color=WorstCode)) +
  facet_grid(rows=vars(WorstCode)) + 
  geom_histogram(aes(y=..density..), alpha=0.5, bins=5) +
  theme_test() +
  scale_color_manual(values=c("grey", "#afb860"), name="") +
  scale_fill_manual(values=c("grey", "#afb860"), name="") +
  labs(x="Number of split codon blocks", y="Density") +
  theme(strip.background = element_blank(), strip.text.y = element_blank()
)
```

## Number of stop codons
```{r}
t.test(numStop ~ WorstCode, res_gb1)


p_numStop = ggplot(res_gb1, aes(x=numStop, fill=WorstCode, color=WorstCode)) +
  facet_grid(rows=vars(WorstCode)) + 
  geom_histogram(aes(y=..density..), alpha=0.5, bins=8) +
  theme_test() +
  scale_color_manual(values=c("grey", "#afb860"), name="") +
  scale_fill_manual(values=c("grey", "#afb860"), name="") +
  labs(x="Number of stop codons", y="Density") +
  theme(strip.background = element_blank(), strip.text.y = element_blank()) +
  scale_x_continuous(breaks=2:9)
```

## The final plot
```{r}
ggarrange(p_venn, 
          ggarrange(p_rob, p_numSplit, p_numStop,
          ncol=3, labels=c("B", "C", "D"), common.legend=TRUE, legend="bottom"),
          ncol=2, labels=c("A", ""), widths=c(1.5,3))
```

## Badly-performing codes
```{r}
# code with the max number of stop codons
bad_codes = res_gb1[res_gb1$WorstCode=="Consistently low-ranking",]
bad_codes[bad_codes$numStop==max(bad_codes$numStop),1:10]

# lowest robustness
bad_codes[bad_codes$Robustness==min(bad_codes$Robustness),1:10]

# max number of split codon blocks
tmp = bad_codes[bad_codes$numSplitBlocks==max(bad_codes$numSplitBlocks),]
# of those, min robustness
tmp[tmp$Robustness==min(tmp$Robustness),1:10]

# Only UAG changes
bad_codes[bad_codes$Block1=="-" & bad_codes$Block3=="X" & bad_codes$Block4=="-" & bad_codes$Block2!="-" ,1:10]
```

## Smaller proportion of worst codes
```{r}
# 5%
res_gb1 = res_gb1[order(res_gb1$MeanFitness),]
worstCodes_gb1 = res_gb1[1:ceiling(0.05*194481),]
res_pare2 = res_pare2[order(res_pare2$MeanFitness),]
worstCodes_pare2 = res_pare2[1:ceiling(0.05*194481),]
res_pare3 = res_pare3[order(res_pare3$MeanFitness),]
worstCodes_pare3 = res_pare3[1:ceiling(0.05*194481),]

worstCodes = intersect(intersect(worstCodes_gb1$Id, worstCodes_pare2$Id), worstCodes_pare3$Id)
length(worstCodes)

# 2%
res_gb1 = res_gb1[order(res_gb1$MeanFitness),]
worstCodes_gb1 = res_gb1[1:ceiling(0.02*194481),]
res_pare2 = res_pare2[order(res_pare2$MeanFitness),]
worstCodes_pare2 = res_pare2[1:ceiling(0.02*194481),]
res_pare3 = res_pare3[order(res_pare3$MeanFitness),]
worstCodes_pare3 = res_pare3[1:ceiling(0.02*194481),]

worstCodes = intersect(intersect(worstCodes_gb1$Id, worstCodes_pare2$Id), worstCodes_pare3$Id)
length(worstCodes)
res_gb1[res_gb1$Id==worstCodes[1],1:10]
```

# Comparison with Pines and Calles codes
How many of the Ostrov codes reach higher fitness than the Calles and Pines codes?

## GB1
```{r}
res_pines_calles = read.delim("../../GB1/03_greedy_walks/pines_calles_codes/output/results", header=FALSE)
for (i in 1:8) {
  print(res_pines_calles[i,1])
  print(sum(res_gb1$MeanFitness>=res_pines_calles[i,2])/194481)
}
```

## ParE2
```{r}
res_pines_calles = read.delim("../../ParE2/03_greedy_walks/pines_calles_codes/output/results", header=FALSE)
for (i in 1:8) {
  print(res_pines_calles[i,1])
  print(sum(res_pare2$MeanFitness>=res_pines_calles[i,2])/194481)
}
```


## ParE3
```{r}
res_pines_calles = read.delim("../../ParE3/03_greedy_walks/pines_calles_codes/output/results", header=FALSE)
for (i in 1:8) {
  print(res_pines_calles[i,1])
  print(sum(res_pare3$MeanFitness>=res_pines_calles[i,2])/194481)
}
```